### 2014-02-28 paper/analysis_lenin_imperialism_the_highest_.tex
add tex extension to analysis tables. add analysis tables also

    \begin{center}
      \begin{tabular}{ | l | l | l |}
        \hline
        Choosen Sensens & Closest Top Sense & similarity  \\ \hline
\multicolumn{3}{|c|}{Lenin: Imperialism, the Highes - path similarity} \\ \hline
capitalism.n.01 & capitalism.n.01 & 1.0 \\ \hline
proletarian.n.01 & great.n.01 & 0.2 \\ \hline
hair.n.01 & universe.n.01 & 0.2 \\ \hline
money.n.01 & penny.n.02 & 0.166666666667 \\ \hline
practice.n.03 & monopoly.n.01 & 0.166666666667 \\ \hline
toilet.n.04 & monopoly.n.01 & 0.142857142857 \\ \hline
war.n.01 & monopoly.n.01 & 0.142857142857 \\ \hline
revolution.n.01 & monopoly.n.01 & 0.125 \\ \hline
theory.n.01 & imperialism.n.01 & 0.125 \\ \hline
television receiver.n.01 & product.n.02 & 0.111111111111 \\ \hline
car.n.01 & product.n.02 & 0.111111111111 \\ \hline
philosophy.n.02 & policy.n.01 & 0.111111111111 \\ \hline
materialism.n.02 & policy.n.01 & 0.111111111111 \\ \hline
strategy.n.02 & policy.n.01 & 0.111111111111 \\ \hline
idealism.n.01 & policy.n.01 & 0.111111111111 \\ \hline
imperialism.n.02 & policy.n.01 & 0.1 \\ \hline
lion.n.01 & great.n.01 & 0.0769230769231 \\ \hline
sugar.n.01 & universe.n.01 & 0.0769230769231 \\ \hline
\multicolumn{3}{|c|}{Lenin: Imperialism, the Highes - lch similarity} \\ \hline
capitalism.n.01 & capitalism.n.01 & 3.63758615973 \\ \hline
proletarian.n.01 & great.n.01 & 2.02814824729 \\ \hline
hair.n.01 & universe.n.01 & 2.02814824729 \\ \hline
money.n.01 & penny.n.02 & 1.8458266905 \\ \hline
practice.n.03 & monopoly.n.01 & 1.8458266905 \\ \hline
toilet.n.04 & monopoly.n.01 & 1.69167601067 \\ \hline
war.n.01 & monopoly.n.01 & 1.69167601067 \\ \hline
revolution.n.01 & monopoly.n.01 & 1.55814461805 \\ \hline
theory.n.01 & imperialism.n.01 & 1.55814461805 \\ \hline
television receiver.n.01 & product.n.02 & 1.44036158239 \\ \hline
car.n.01 & product.n.02 & 1.44036158239 \\ \hline
philosophy.n.02 & policy.n.01 & 1.44036158239 \\ \hline
materialism.n.02 & policy.n.01 & 1.44036158239 \\ \hline
strategy.n.02 & policy.n.01 & 1.44036158239 \\ \hline
idealism.n.01 & policy.n.01 & 1.44036158239 \\ \hline
imperialism.n.02 & policy.n.01 & 1.33500106673 \\ \hline
lion.n.01 & great.n.01 & 1.07263680226 \\ \hline
sugar.n.01 & universe.n.01 & 1.07263680226 \\ \hline
\multicolumn{3}{|c|}{Lenin: Imperialism, the Highes - res similarity} \\ \hline
capitalism.n.01 & capitalism.n.01 & 11.6930111908 \\ \hline
money.n.01 & penny.n.02 & 6.2549318819 \\ \hline
theory.n.01 & imperialism.n.01 & 5.678990458 \\ \hline
hair.n.01 & universe.n.01 & 4.94037581121 \\ \hline
philosophy.n.02 & policy.n.01 & 3.29228159341 \\ \hline
materialism.n.02 & policy.n.01 & 3.29228159341 \\ \hline
strategy.n.02 & policy.n.01 & 3.29228159341 \\ \hline
idealism.n.01 & policy.n.01 & 3.29228159341 \\ \hline
toilet.n.04 & monopoly.n.01 & 3.14352023311 \\ \hline
imperialism.n.02 & imperialism.n.01 & 2.64452085446 \\ \hline
proletarian.n.01 & great.n.01 & 2.33354524374 \\ \hline
war.n.01 & monopoly.n.01 & 2.33223775508 \\ \hline
practice.n.03 & monopoly.n.01 & 2.33223775508 \\ \hline
television receiver.n.01 & product.n.02 & 2.30584873145 \\ \hline
car.n.01 & product.n.02 & 2.30584873145 \\ \hline
lion.n.01 & great.n.01 & 2.22415047123 \\ \hline
revolution.n.01 & monopoly.n.01 & 2.03632771779 \\ \hline
sugar.n.01 & universe.n.01 & 0.801759114954 \\ \hline
\multicolumn{3}{|c|}{Lenin: Imperialism, the Highes - jcn similarity} \\ \hline
capitalism.n.01 & capitalism.n.01 & 1e+300 \\ \hline
money.n.01 & penny.n.02 & 0.220543285308 \\ \hline
hair.n.01 & universe.n.01 & 0.185089245333 \\ \hline
car.n.01 & product.n.02 & 0.118562212358 \\ \hline
philosophy.n.02 & policy.n.01 & 0.0941972384174 \\ \hline
television receiver.n.01 & product.n.02 & 0.0922281130906 \\ \hline
theory.n.01 & imperialism.n.01 & 0.0849319922588 \\ \hline
proletarian.n.01 & product.n.02 & 0.0827408100653 \\ \hline
war.n.01 & monopoly.n.01 & 0.0769380100151 \\ \hline
lion.n.01 & product.n.02 & 0.0745010379336 \\ \hline
strategy.n.02 & policy.n.01 & 0.0729992753311 \\ \hline
practice.n.03 & monopoly.n.01 & 0.0696858132394 \\ \hline
toilet.n.04 & monopoly.n.01 & 0.0675356263653 \\ \hline
materialism.n.02 & policy.n.01 & 0.0673677474158 \\ \hline
sugar.n.01 & product.n.02 & 0.0665737679313 \\ \hline
revolution.n.01 & one.n.01 & 0.0654799767733 \\ \hline
idealism.n.01 & policy.n.01 & 0.063730003856 \\ \hline
imperialism.n.02 & policy.n.01 & 0.0574970955164 \\ \hline
\multicolumn{3}{|c|}{Lenin: Imperialism, the Highes - lin similarity} \\ \hline
capitalism.n.01 & capitalism.n.01 & 1.0 \\ \hline
money.n.01 & penny.n.02 & 0.733969426851 \\ \hline
hair.n.01 & universe.n.01 & 0.646495819603 \\ \hline
theory.n.01 & imperialism.n.01 & 0.491005027416 \\ \hline
philosophy.n.02 & policy.n.01 & 0.382810406388 \\ \hline
car.n.01 & product.n.02 & 0.353492745777 \\ \hline
strategy.n.02 & policy.n.01 & 0.324629309412 \\ \hline
materialism.n.02 & policy.n.01 & 0.307281190083 \\ \hline
television receiver.n.01 & product.n.02 & 0.298407179857 \\ \hline
toilet.n.04 & monopoly.n.01 & 0.298048188669 \\ \hline
idealism.n.01 & policy.n.01 & 0.295593207225 \\ \hline
war.n.01 & monopoly.n.01 & 0.264097390196 \\ \hline
proletarian.n.01 & great.n.01 & 0.262634686841 \\ \hline
practice.n.03 & monopoly.n.01 & 0.245310227149 \\ \hline
imperialism.n.02 & policy.n.01 & 0.233190306342 \\ \hline
lion.n.01 & great.n.01 & 0.232810355981 \\ \hline
revolution.n.01 & monopoly.n.01 & 0.205174493173 \\ \hline
sugar.n.01 & product.n.02 & 0.0964554176079 \\ \hline
\end{tabular}
\end{center}