### 2014-02-28 paper/analysis_lenin_materialism_and_empirio-.tex

    \begin{center}
      \begin{tabular}{ | l | l | l |}
        \hline
        Choosen Sensens & Closest Top Sense & similarity  \\ \hline
\multicolumn{3}{|c|}{Lenin: MATERIALISM and EMPIRIO - path similarity} \\ \hline
revolution.n.01 & experience.n.03 & 0.25 \\ \hline
proletarian.n.01 & matter.n.03 & 0.166666666667 \\ \hline
philosophy.n.02 & physics.n.01 & 0.166666666667 \\ \hline
war.n.01 & experience.n.03 & 0.166666666667 \\ \hline
strategy.n.02 & physics.n.01 & 0.166666666667 \\ \hline
practice.n.03 & experience.n.03 & 0.166666666667 \\ \hline
money.n.01 & beginning.n.02 & 0.142857142857 \\ \hline
materialism.n.02 & aim.n.02 & 0.142857142857 \\ \hline
idealism.n.01 & aim.n.02 & 0.142857142857 \\ \hline
imperialism.n.02 & aim.n.02 & 0.125 \\ \hline
sugar.n.01 & matter.n.03 & 0.125 \\ \hline
theory.n.01 & sensation.n.01 & 0.125 \\ \hline
hair.n.01 & space.n.02 & 0.125 \\ \hline
capitalism.n.01 & time.n.05 & 0.125 \\ \hline
toilet.n.04 & experience.n.03 & 0.111111111111 \\ \hline
television receiver.n.01 & space.n.02 & 0.0909090909091 \\ \hline
car.n.01 & space.n.02 & 0.0909090909091 \\ \hline
lion.n.01 & kant.n.01 & 0.0666666666667 \\ \hline
\multicolumn{3}{|c|}{Lenin: MATERIALISM and EMPIRIO - lch similarity} \\ \hline
revolution.n.01 & experience.n.03 & 2.25129179861 \\ \hline
proletarian.n.01 & matter.n.03 & 1.8458266905 \\ \hline
philosophy.n.02 & physics.n.01 & 1.8458266905 \\ \hline
war.n.01 & experience.n.03 & 1.8458266905 \\ \hline
strategy.n.02 & physics.n.01 & 1.8458266905 \\ \hline
practice.n.03 & experience.n.03 & 1.8458266905 \\ \hline
money.n.01 & beginning.n.02 & 1.69167601067 \\ \hline
materialism.n.02 & aim.n.02 & 1.69167601067 \\ \hline
idealism.n.01 & aim.n.02 & 1.69167601067 \\ \hline
imperialism.n.02 & aim.n.02 & 1.55814461805 \\ \hline
sugar.n.01 & matter.n.03 & 1.55814461805 \\ \hline
theory.n.01 & sensation.n.01 & 1.55814461805 \\ \hline
hair.n.01 & space.n.02 & 1.55814461805 \\ \hline
capitalism.n.01 & time.n.05 & 1.55814461805 \\ \hline
toilet.n.04 & experience.n.03 & 1.44036158239 \\ \hline
television receiver.n.01 & space.n.02 & 1.23969088693 \\ \hline
car.n.01 & space.n.02 & 1.23969088693 \\ \hline
lion.n.01 & kant.n.01 & 0.929535958624 \\ \hline
\multicolumn{3}{|c|}{Lenin: MATERIALISM and EMPIRIO - res similarity} \\ \hline
philosophy.n.02 & physics.n.01 & 5.58877614988 \\ \hline
strategy.n.02 & physics.n.01 & 5.58877614988 \\ \hline
theory.n.01 & sensation.n.01 & 4.38484522248 \\ \hline
revolution.n.01 & experience.n.03 & 4.05348916032 \\ \hline
sugar.n.01 & matter.n.03 & 3.3367059068 \\ \hline
materialism.n.02 & physics.n.01 & 3.29228159341 \\ \hline
idealism.n.01 & physics.n.01 & 3.29228159341 \\ \hline
capitalism.n.01 & sociable.n.01 & 2.85529368615 \\ \hline
money.n.01 & beginning.n.02 & 2.69110280536 \\ \hline
imperialism.n.02 & physics.n.01 & 2.64452085446 \\ \hline
proletarian.n.01 & kant.n.01 & 2.33354524374 \\ \hline
toilet.n.04 & motion.n.03 & 2.33223775508 \\ \hline
war.n.01 & motion.n.03 & 2.33223775508 \\ \hline
practice.n.03 & motion.n.03 & 2.33223775508 \\ \hline
television receiver.n.01 & book.n.01 & 2.30584873145 \\ \hline
car.n.01 & book.n.01 & 2.30584873145 \\ \hline
lion.n.01 & kant.n.01 & 2.22415047123 \\ \hline
hair.n.01 & kant.n.01 & 1.53183374322 \\ \hline
\multicolumn{3}{|c|}{Lenin: MATERIALISM and EMPIRIO - jcn similarity} \\ \hline
philosophy.n.02 & physics.n.01 & 0.145932703696 \\ \hline
theory.n.01 & sensation.n.01 & 0.128286683846 \\ \hline
sugar.n.01 & matter.n.03 & 0.127664075736 \\ \hline
revolution.n.01 & experience.n.03 & 0.111569245855 \\ \hline
money.n.01 & moment.n.01 & 0.107804699133 \\ \hline
car.n.01 & matter.n.03 & 0.107243331568 \\ \hline
hair.n.01 & matter.n.03 & 0.104557075904 \\ \hline
strategy.n.02 & physics.n.01 & 0.100652051627 \\ \hline
war.n.01 & motion.n.03 & 0.100254651557 \\ \hline
practice.n.03 & motion.n.03 & 0.0882827134582 \\ \hline
proletarian.n.01 & matter.n.03 & 0.0875035868175 \\ \hline
television receiver.n.01 & matter.n.03 & 0.085230569567 \\ \hline
lion.n.01 & matter.n.03 & 0.0783404338338 \\ \hline
materialism.n.02 & aim.n.02 & 0.0778269817906 \\ \hline
toilet.n.04 & motion.n.03 & 0.0745896410117 \\ \hline
idealism.n.01 & aim.n.02 & 0.0730123547216 \\ \hline
capitalism.n.01 & sociable.n.01 & 0.0678992095349 \\ \hline
imperialism.n.02 & aim.n.02 & 0.0649464570442 \\ \hline
\multicolumn{3}{|c|}{Lenin: MATERIALISM and EMPIRIO - lin similarity} \\ \hline
philosophy.n.02 & physics.n.01 & 0.619941000616 \\ \hline
strategy.n.02 & physics.n.01 & 0.529421413479 \\ \hline
theory.n.01 & sensation.n.01 & 0.529419405948 \\ \hline
revolution.n.01 & experience.n.03 & 0.474924895953 \\ \hline
sugar.n.01 & matter.n.03 & 0.460030062094 \\ \hline
money.n.01 & moment.n.01 & 0.367179548114 \\ \hline
materialism.n.02 & aim.n.02 & 0.33882403792 \\ \hline
idealism.n.01 & aim.n.02 & 0.324668589607 \\ \hline
war.n.01 & motion.n.03 & 0.318631846502 \\ \hline
car.n.01 & book.n.01 & 0.308668677605 \\ \hline
practice.n.03 & motion.n.03 & 0.291680639247 \\ \hline
capitalism.n.01 & sociable.n.01 & 0.27940619133 \\ \hline
television receiver.n.01 & book.n.01 & 0.265820778874 \\ \hline
toilet.n.04 & motion.n.03 & 0.258117063881 \\ \hline
imperialism.n.02 & aim.n.02 & 0.255677978749 \\ \hline
hair.n.01 & book.n.01 & 0.20182033649 \\ \hline
proletarian.n.01 & kant.n.01 & 0.198875753057 \\ \hline
lion.n.01 & kant.n.01 & 0.179337574844 \\ \hline
\end{tabular}
\end{center}