### 2014-02-28 paper/analysis_lenin_the_agrarian_programme_o.tex

    \begin{center}
      \begin{tabular}{ | l | l | l |}
        \hline
        Choosen Sensens & Closest Top Sense & similarity  \\ \hline
\multicolumn{3}{|c|}{Lenin: The Agrarian Programme  - path similarity} \\ \hline
revolution.n.01 & revolution.n.01 & 1.0 \\ \hline
proletarian.n.01 & peasant.n.01 & 0.25 \\ \hline
money.n.01 & bill.n.03 & 0.2 \\ \hline
war.n.01 & nationalization.n.03 & 0.2 \\ \hline
practice.n.03 & nationalization.n.03 & 0.166666666667 \\ \hline
capitalism.n.01 & people.n.01 & 0.166666666667 \\ \hline
hair.n.01 & land.n.04 & 0.142857142857 \\ \hline
toilet.n.04 & nationalization.n.03 & 0.111111111111 \\ \hline
philosophy.n.02 & absolute.n.01 & 0.111111111111 \\ \hline
materialism.n.02 & absolute.n.01 & 0.111111111111 \\ \hline
strategy.n.02 & absolute.n.01 & 0.111111111111 \\ \hline
idealism.n.01 & absolute.n.01 & 0.111111111111 \\ \hline
imperialism.n.02 & people.n.01 & 0.111111111111 \\ \hline
television receiver.n.01 & land.n.04 & 0.1 \\ \hline
car.n.01 & land.n.04 & 0.1 \\ \hline
theory.n.01 & people.n.01 & 0.1 \\ \hline
sugar.n.01 & land.n.04 & 0.0909090909091 \\ \hline
lion.n.01 & land.n.04 & 0.0714285714286 \\ \hline
\multicolumn{3}{|c|}{Lenin: The Agrarian Programme  - lch similarity} \\ \hline
revolution.n.01 & revolution.n.01 & 3.63758615973 \\ \hline
proletarian.n.01 & peasant.n.01 & 2.25129179861 \\ \hline
money.n.01 & bill.n.03 & 2.02814824729 \\ \hline
war.n.01 & nationalization.n.03 & 2.02814824729 \\ \hline
practice.n.03 & nationalization.n.03 & 1.8458266905 \\ \hline
capitalism.n.01 & people.n.01 & 1.8458266905 \\ \hline
hair.n.01 & land.n.04 & 1.69167601067 \\ \hline
toilet.n.04 & nationalization.n.03 & 1.44036158239 \\ \hline
philosophy.n.02 & absolute.n.01 & 1.44036158239 \\ \hline
materialism.n.02 & absolute.n.01 & 1.44036158239 \\ \hline
strategy.n.02 & absolute.n.01 & 1.44036158239 \\ \hline
idealism.n.01 & absolute.n.01 & 1.44036158239 \\ \hline
imperialism.n.02 & people.n.01 & 1.44036158239 \\ \hline
television receiver.n.01 & land.n.04 & 1.33500106673 \\ \hline
car.n.01 & land.n.04 & 1.33500106673 \\ \hline
theory.n.01 & people.n.01 & 1.33500106673 \\ \hline
sugar.n.01 & land.n.04 & 1.23969088693 \\ \hline
lion.n.01 & land.n.04 & 0.998528830111 \\ \hline
\multicolumn{3}{|c|}{Lenin: The Agrarian Programme  - res similarity} \\ \hline
revolution.n.01 & revolution.n.01 & 10.3880624692 \\ \hline
proletarian.n.01 & peasant.n.01 & 9.00601439892 \\ \hline
money.n.01 & bill.n.03 & 6.2549318819 \\ \hline
war.n.01 & nationalization.n.03 & 4.52385364841 \\ \hline
philosophy.n.02 & absolute.n.01 & 3.29228159341 \\ \hline
materialism.n.02 & absolute.n.01 & 3.29228159341 \\ \hline
strategy.n.02 & absolute.n.01 & 3.29228159341 \\ \hline
idealism.n.01 & absolute.n.01 & 3.29228159341 \\ \hline
capitalism.n.01 & people.n.01 & 2.85529368615 \\ \hline
imperialism.n.02 & absolute.n.01 & 2.64452085446 \\ \hline
theory.n.01 & absolute.n.01 & 2.64452085446 \\ \hline
toilet.n.04 & nationalization.n.03 & 2.33223775508 \\ \hline
practice.n.03 & nationalization.n.03 & 2.33223775508 \\ \hline
lion.n.01 & peasant.n.01 & 2.22415047123 \\ \hline
television receiver.n.01 & peasant.n.01 & 1.53183374322 \\ \hline
car.n.01 & peasant.n.01 & 1.53183374322 \\ \hline
hair.n.01 & peasant.n.01 & 1.53183374322 \\ \hline
sugar.n.01 & land.n.04 & 0.801759114954 \\ \hline
\multicolumn{3}{|c|}{Lenin: The Agrarian Programme  - jcn similarity} \\ \hline
revolution.n.01 & revolution.n.01 & 1e+300 \\ \hline
proletarian.n.01 & peasant.n.01 & 0.363995193783 \\ \hline
money.n.01 & bill.n.03 & 0.30619633675 \\ \hline
war.n.01 & farming.n.01 & 0.102813269009 \\ \hline
car.n.01 & land.n.04 & 0.0903577501322 \\ \hline
hair.n.01 & land.n.04 & 0.0884432572523 \\ \hline
capitalism.n.01 & people.n.01 & 0.0880782305893 \\ \hline
philosophy.n.02 & people.n.01 & 0.0801894085342 \\ \hline
television receiver.n.01 & land.n.04 & 0.0742092466431 \\ \hline
theory.n.01 & people.n.01 & 0.0738644334041 \\ \hline
practice.n.03 & people.n.01 & 0.0728246934249 \\ \hline
lion.n.01 & land.n.04 & 0.0689306688111 \\ \hline
strategy.n.02 & people.n.01 & 0.0642953740963 \\ \hline
sugar.n.01 & land.n.04 & 0.0640122327597 \\ \hline
toilet.n.04 & people.n.01 & 0.0632469161145 \\ \hline
materialism.n.02 & people.n.01 & 0.0598861536358 \\ \hline
idealism.n.01 & people.n.01 & 0.056994188175 \\ \hline
imperialism.n.02 & people.n.01 & 0.055706849963 \\ \hline
\multicolumn{3}{|c|}{Lenin: The Agrarian Programme  - lin similarity} \\ \hline
revolution.n.01 & revolution.n.01 & 1.0 \\ \hline
proletarian.n.01 & peasant.n.01 & 0.867659956556 \\ \hline
money.n.01 & bill.n.03 & 0.792980998294 \\ \hline
war.n.01 & farming.n.01 & 0.481925511595 \\ \hline
philosophy.n.02 & absolute.n.01 & 0.341316252969 \\ \hline
capitalism.n.01 & people.n.01 & 0.334654457361 \\ \hline
strategy.n.02 & absolute.n.01 & 0.294289762784 \\ \hline
materialism.n.02 & absolute.n.01 & 0.279961229282 \\ \hline
idealism.n.01 & absolute.n.01 & 0.270226260066 \\ \hline
theory.n.01 & absolute.n.01 & 0.259782259686 \\ \hline
practice.n.03 & restoration.n.02 & 0.240021453557 \\ \hline
toilet.n.04 & restoration.n.02 & 0.216821056563 \\ \hline
imperialism.n.02 & absolute.n.01 & 0.213506139286 \\ \hline
lion.n.01 & peasant.n.01 & 0.20131688425 \\ \hline
car.n.01 & land.n.04 & 0.18905393848 \\ \hline
hair.n.01 & land.n.04 & 0.185792500123 \\ \hline
television receiver.n.01 & land.n.04 & 0.160696186575 \\ \hline
sugar.n.01 & land.n.04 & 0.0930896185493 \\ \hline
\end{tabular}
\end{center}