### 2014-02-28 paper/analysis_lenin_what_is_to_be_done?.tex

    \begin{center}
      \begin{tabular}{ | l | l | l |}
        \hline
        Choosen Sensens & Closest Top Sense & similarity  \\ \hline
\multicolumn{3}{|c|}{Lenin: What Is To Be Done? - path similarity} \\ \hline
imperialism.n.02 & political orientation.n.01 & 0.5 \\ \hline
proletarian.n.01 & worker.n.01 & 0.25 \\ \hline
war.n.01 & conflict.n.01 & 0.25 \\ \hline
practice.n.03 & conflict.n.01 & 0.2 \\ \hline
revolution.n.01 & conflict.n.01 & 0.166666666667 \\ \hline
toilet.n.04 & trade.n.02 & 0.142857142857 \\ \hline
money.n.01 & mass.n.01 & 0.142857142857 \\ \hline
capitalism.n.01 & organization.n.01 & 0.142857142857 \\ \hline
hair.n.01 & working.n.01 & 0.125 \\ \hline
television receiver.n.01 & working.n.01 & 0.111111111111 \\ \hline
car.n.01 & working.n.01 & 0.111111111111 \\ \hline
philosophy.n.02 & political orientation.n.01 & 0.111111111111 \\ \hline
materialism.n.02 & political orientation.n.01 & 0.111111111111 \\ \hline
strategy.n.02 & political orientation.n.01 & 0.111111111111 \\ \hline
idealism.n.01 & political orientation.n.01 & 0.111111111111 \\ \hline
theory.n.01 & political orientation.n.01 & 0.111111111111 \\ \hline
sugar.n.01 & local anesthetic.n.01 & 0.0909090909091 \\ \hline
lion.n.01 & worker.n.01 & 0.0833333333333 \\ \hline
\multicolumn{3}{|c|}{Lenin: What Is To Be Done? - lch similarity} \\ \hline
imperialism.n.02 & political orientation.n.01 & 2.94443897917 \\ \hline
proletarian.n.01 & worker.n.01 & 2.25129179861 \\ \hline
war.n.01 & conflict.n.01 & 2.25129179861 \\ \hline
practice.n.03 & conflict.n.01 & 2.02814824729 \\ \hline
revolution.n.01 & conflict.n.01 & 1.8458266905 \\ \hline
toilet.n.04 & trade.n.02 & 1.69167601067 \\ \hline
money.n.01 & mass.n.01 & 1.69167601067 \\ \hline
capitalism.n.01 & organization.n.01 & 1.69167601067 \\ \hline
hair.n.01 & working.n.01 & 1.55814461805 \\ \hline
television receiver.n.01 & working.n.01 & 1.44036158239 \\ \hline
car.n.01 & working.n.01 & 1.44036158239 \\ \hline
philosophy.n.02 & political orientation.n.01 & 1.44036158239 \\ \hline
materialism.n.02 & political orientation.n.01 & 1.44036158239 \\ \hline
strategy.n.02 & political orientation.n.01 & 1.44036158239 \\ \hline
idealism.n.01 & political orientation.n.01 & 1.44036158239 \\ \hline
theory.n.01 & political orientation.n.01 & 1.44036158239 \\ \hline
sugar.n.01 & local anesthetic.n.01 & 1.23969088693 \\ \hline
lion.n.01 & worker.n.01 & 1.15267950994 \\ \hline
\multicolumn{3}{|c|}{Lenin: What Is To Be Done? - res similarity} \\ \hline
imperialism.n.02 & political orientation.n.01 & 8.60196873746 \\ \hline
sugar.n.01 & local anesthetic.n.01 & 4.61519673162 \\ \hline
war.n.01 & conflict.n.01 & 4.52385364841 \\ \hline
television receiver.n.01 & newspaper.n.01 & 3.25767902589 \\ \hline
car.n.01 & newspaper.n.01 & 3.25767902589 \\ \hline
toilet.n.04 & trade.n.02 & 3.14352023311 \\ \hline
capitalism.n.01 & organization.n.01 & 2.85529368615 \\ \hline
money.n.01 & mass.n.01 & 2.69110280536 \\ \hline
philosophy.n.02 & political orientation.n.01 & 2.64452085446 \\ \hline
materialism.n.02 & political orientation.n.01 & 2.64452085446 \\ \hline
strategy.n.02 & political orientation.n.01 & 2.64452085446 \\ \hline
idealism.n.01 & political orientation.n.01 & 2.64452085446 \\ \hline
theory.n.01 & political orientation.n.01 & 2.64452085446 \\ \hline
proletarian.n.01 & worker.n.01 & 2.33354524374 \\ \hline
practice.n.03 & trade.n.02 & 2.33223775508 \\ \hline
lion.n.01 & worker.n.01 & 2.22415047123 \\ \hline
revolution.n.01 & trade.n.02 & 2.03632771779 \\ \hline
hair.n.01 & worker.n.01 & 1.53183374322 \\ \hline
\multicolumn{3}{|c|}{Lenin: What Is To Be Done? - jcn similarity} \\ \hline
imperialism.n.02 & political orientation.n.01 & 0.193405491818 \\ \hline
war.n.01 & conflict.n.01 & 0.152829762599 \\ \hline
car.n.01 & worker.n.01 & 0.110550521246 \\ \hline
hair.n.01 & worker.n.01 & 0.107698235698 \\ \hline
proletarian.n.01 & worker.n.01 & 0.104758893715 \\ \hline
capitalism.n.01 & organization.n.01 & 0.101121795595 \\ \hline
sugar.n.01 & local anesthetic.n.01 & 0.0929714058017 \\ \hline
money.n.01 & organization.n.01 & 0.0929667353193 \\ \hline
philosophy.n.02 & organization.n.01 & 0.0908595602215 \\ \hline
lion.n.01 & worker.n.01 & 0.0900802228812 \\ \hline
practice.n.03 & motion.n.03 & 0.0882827134582 \\ \hline
television receiver.n.01 & worker.n.01 & 0.0873062903722 \\ \hline
theory.n.01 & organization.n.01 & 0.0828237113739 \\ \hline
revolution.n.01 & motion.n.03 & 0.0784056470694 \\ \hline
toilet.n.04 & motion.n.03 & 0.0745896410117 \\ \hline
strategy.n.02 & organization.n.01 & 0.0709786674118 \\ \hline
materialism.n.02 & organization.n.01 & 0.0656431911358 \\ \hline
idealism.n.01 & organization.n.01 & 0.0621845289629 \\ \hline
\multicolumn{3}{|c|}{Lenin: What Is To Be Done? - lin similarity} \\ \hline
imperialism.n.02 & political orientation.n.01 & 0.768910941363 \\ \hline
war.n.01 & conflict.n.01 & 0.580318438615 \\ \hline
sugar.n.01 & local anesthetic.n.01 & 0.461833980698 \\ \hline
car.n.01 & newspaper.n.01 & 0.411156952767 \\ \hline
capitalism.n.01 & organization.n.01 & 0.366071452787 \\ \hline
television receiver.n.01 & newspaper.n.01 & 0.356914166707 \\ \hline
money.n.01 & mass.n.01 & 0.331362497131 \\ \hline
proletarian.n.01 & worker.n.01 & 0.328371898491 \\ \hline
philosophy.n.02 & political orientation.n.01 & 0.31307622756 \\ \hline
toilet.n.04 & trade.n.02 & 0.312879932381 \\ \hline
theory.n.01 & political orientation.n.01 & 0.294463460062 \\ \hline
practice.n.03 & motion.n.03 & 0.291680639247 \\ \hline
lion.n.01 & worker.n.01 & 0.286073258452 \\ \hline
strategy.n.02 & political orientation.n.01 & 0.264762834975 \\ \hline
materialism.n.02 & political orientation.n.01 & 0.25040844726 \\ \hline
hair.n.01 & worker.n.01 & 0.248092928538 \\ \hline
revolution.n.01 & motion.n.03 & 0.242033306584 \\ \hline
idealism.n.01 & political orientation.n.01 & 0.240750703939 \\ \hline
\end{tabular}
\end{center}