### 2014-02-28 paper/analysis_the_state_and_revolution.tex

    \begin{center}
      \begin{tabular}{ | l | l | l |}
        \hline
        Choosen Sensens & Closest Top Sense & similarity  \\ \hline
\multicolumn{3}{|c|}{The State and Revolution - path similarity} \\ \hline
revolution.n.01 & revolution.n.01 & 1.0 \\ \hline
imperialism.n.02 & communism.n.02 & 0.333333333333 \\ \hline
proletarian.n.01 & peer.n.01 & 0.25 \\ \hline
practice.n.03 & work.n.01 & 0.2 \\ \hline
toilet.n.04 & work.n.01 & 0.166666666667 \\ \hline
war.n.01 & work.n.01 & 0.166666666667 \\ \hline
money.n.01 & phase.n.01 & 0.125 \\ \hline
philosophy.n.02 & law.n.04 & 0.125 \\ \hline
materialism.n.02 & law.n.04 & 0.125 \\ \hline
strategy.n.02 & law.n.04 & 0.125 \\ \hline
idealism.n.01 & law.n.04 & 0.125 \\ \hline
capitalism.n.01 & democracy.n.02 & 0.125 \\ \hline
hair.n.01 & peer.n.01 & 0.111111111111 \\ \hline
theory.n.01 & experience.n.03 & 0.1 \\ \hline
television receiver.n.01 & peer.n.01 & 0.0833333333333 \\ \hline
car.n.01 & peer.n.01 & 0.0833333333333 \\ \hline
lion.n.01 & peer.n.01 & 0.0833333333333 \\ \hline
sugar.n.01 & peer.n.01 & 0.0833333333333 \\ \hline
\multicolumn{3}{|c|}{The State and Revolution - lch similarity} \\ \hline
revolution.n.01 & revolution.n.01 & 3.63758615973 \\ \hline
imperialism.n.02 & communism.n.02 & 2.53897387106 \\ \hline
proletarian.n.01 & peer.n.01 & 2.25129179861 \\ \hline
practice.n.03 & work.n.01 & 2.02814824729 \\ \hline
toilet.n.04 & work.n.01 & 1.8458266905 \\ \hline
war.n.01 & work.n.01 & 1.8458266905 \\ \hline
money.n.01 & phase.n.01 & 1.55814461805 \\ \hline
philosophy.n.02 & law.n.04 & 1.55814461805 \\ \hline
materialism.n.02 & law.n.04 & 1.55814461805 \\ \hline
strategy.n.02 & law.n.04 & 1.55814461805 \\ \hline
idealism.n.01 & law.n.04 & 1.55814461805 \\ \hline
capitalism.n.01 & democracy.n.02 & 1.55814461805 \\ \hline
hair.n.01 & peer.n.01 & 1.44036158239 \\ \hline
theory.n.01 & experience.n.03 & 1.33500106673 \\ \hline
television receiver.n.01 & peer.n.01 & 1.15267950994 \\ \hline
car.n.01 & peer.n.01 & 1.15267950994 \\ \hline
lion.n.01 & peer.n.01 & 1.15267950994 \\ \hline
sugar.n.01 & peer.n.01 & 1.15267950994 \\ \hline
\multicolumn{3}{|c|}{The State and Revolution - res similarity} \\ \hline
revolution.n.01 & revolution.n.01 & 10.3880624692 \\ \hline
imperialism.n.02 & communism.n.02 & 8.60196873746 \\ \hline
philosophy.n.02 & law.n.04 & 3.29228159341 \\ \hline
materialism.n.02 & law.n.04 & 3.29228159341 \\ \hline
strategy.n.02 & law.n.04 & 3.29228159341 \\ \hline
idealism.n.01 & law.n.04 & 3.29228159341 \\ \hline
toilet.n.04 & work.n.01 & 3.14352023311 \\ \hline
capitalism.n.01 & club.n.02 & 2.85529368615 \\ \hline
money.n.01 & phase.n.01 & 2.69110280536 \\ \hline
theory.n.01 & law.n.04 & 2.64452085446 \\ \hline
proletarian.n.01 & engels.n.01 & 2.33354524374 \\ \hline
war.n.01 & work.n.01 & 2.33223775508 \\ \hline
practice.n.03 & work.n.01 & 2.33223775508 \\ \hline
lion.n.01 & engels.n.01 & 2.22415047123 \\ \hline
television receiver.n.01 & engels.n.01 & 1.53183374322 \\ \hline
car.n.01 & engels.n.01 & 1.53183374322 \\ \hline
hair.n.01 & engels.n.01 & 1.53183374322 \\ \hline
sugar.n.01 & engels.n.01 & 0.801759114954 \\ \hline
\multicolumn{3}{|c|}{The State and Revolution - jcn similarity} \\ \hline
revolution.n.01 & revolution.n.01 & 1e+300 \\ \hline
imperialism.n.02 & communism.n.02 & 0.14975874591 \\ \hline
war.n.01 & work.n.01 & 0.119881984187 \\ \hline
philosophy.n.02 & law.n.04 & 0.103411969458 \\ \hline
practice.n.03 & work.n.01 & 0.103154641506 \\ \hline
toilet.n.04 & work.n.01 & 0.0985118809398 \\ \hline
money.n.01 & phase.n.01 & 0.0916158054473 \\ \hline
theory.n.01 & work.n.01 & 0.0910758552239 \\ \hline
car.n.01 & peer.n.01 & 0.0887294617483 \\ \hline
hair.n.01 & peer.n.01 & 0.0868826421077 \\ \hline
proletarian.n.01 & peer.n.01 & 0.0849595683478 \\ \hline
strategy.n.02 & law.n.04 & 0.07841412673 \\ \hline
lion.n.01 & peer.n.01 & 0.0750424463326 \\ \hline
television receiver.n.01 & peer.n.01 & 0.0731074093034 \\ \hline
capitalism.n.01 & club.n.02 & 0.0729022081709 \\ \hline
materialism.n.02 & law.n.04 & 0.0719531280061 \\ \hline
idealism.n.01 & law.n.04 & 0.0678185221973 \\ \hline
sugar.n.01 & work.n.01 & 0.0625959407953 \\ \hline
\multicolumn{3}{|c|}{The State and Revolution - lin similarity} \\ \hline
revolution.n.01 & revolution.n.01 & 1.0 \\ \hline
imperialism.n.02 & communism.n.02 & 0.720392353354 \\ \hline
philosophy.n.02 & law.n.04 & 0.405088626962 \\ \hline
toilet.n.04 & work.n.01 & 0.382467580972 \\ \hline
war.n.01 & work.n.01 & 0.35863993877 \\ \hline
strategy.n.02 & law.n.04 & 0.340509805473 \\ \hline
money.n.01 & phase.n.01 & 0.330250297796 \\ \hline
practice.n.03 & work.n.01 & 0.324854541172 \\ \hline
materialism.n.02 & law.n.04 & 0.321472637881 \\ \hline
idealism.n.01 & law.n.04 & 0.308702564543 \\ \hline
theory.n.01 & law.n.04 & 0.305328480717 \\ \hline
capitalism.n.01 & club.n.02 & 0.29394209418 \\ \hline
proletarian.n.01 & peer.n.01 & 0.28393127113 \\ \hline
lion.n.01 & peer.n.01 & 0.250268807542 \\ \hline
car.n.01 & peer.n.01 & 0.213736072979 \\ \hline
hair.n.01 & peer.n.01 & 0.210222579299 \\ \hline
television receiver.n.01 & peer.n.01 & 0.182991045418 \\ \hline
sugar.n.01 & peer.n.01 & 0.0895207954547 \\ \hline
\end{tabular}
\end{center}