### 2014-03-01 paper/analysis_lenin_imperialism_the_highest_.tex
add graph that compares each sense against each other. add definition to top senses tables

    \begin{center}
      \begin{tabular}{ | l | l | l | l |}
        \hline
        Choosen Sensens & Closest Top Sense & similarity & Top Sense Gloss \\ \hline
\multicolumn{4}{|c|}{Lenin: Imperialism, the Highes - path similarity} \\ \hline
capitalism.n.01 & capitalism.n.01 & 1.0 & an economic system based on private ownership of capital\\ \hline
proletarian.n.01 & great.n.01 & 0.2 & a person who has achieved distinction and honor in some field\\ \hline
hair.n.01 & universe.n.01 & 0.2 & everything that exists anywhere\\ \hline
money.n.01 & penny.n.02 & 0.166666666667 & a coin worth one-hundredth of the value of the basic unit\\ \hline
practice.n.03 & monopoly.n.01 & 0.166666666667 & (economics) a market in which there are many buyers but only one seller\\ \hline
toilet.n.04 & monopoly.n.01 & 0.142857142857 & (economics) a market in which there are many buyers but only one seller\\ \hline
war.n.01 & monopoly.n.01 & 0.142857142857 & (economics) a market in which there are many buyers but only one seller\\ \hline
revolution.n.01 & monopoly.n.01 & 0.125 & (economics) a market in which there are many buyers but only one seller\\ \hline
theory.n.01 & imperialism.n.01 & 0.125 & a policy of extending your rule over foreign countries\\ \hline
television receiver.n.01 & product.n.02 & 0.111111111111 & an artifact that has been created by someone or some process\\ \hline
car.n.01 & product.n.02 & 0.111111111111 & an artifact that has been created by someone or some process\\ \hline
philosophy.n.02 & policy.n.01 & 0.111111111111 & a plan of action adopted by an individual or social group\\ \hline
materialism.n.02 & policy.n.01 & 0.111111111111 & a plan of action adopted by an individual or social group\\ \hline
strategy.n.02 & policy.n.01 & 0.111111111111 & a plan of action adopted by an individual or social group\\ \hline
idealism.n.01 & policy.n.01 & 0.111111111111 & a plan of action adopted by an individual or social group\\ \hline
imperialism.n.02 & policy.n.01 & 0.1 & a plan of action adopted by an individual or social group\\ \hline
lion.n.01 & great.n.01 & 0.0769230769231 & a person who has achieved distinction and honor in some field\\ \hline
sugar.n.01 & universe.n.01 & 0.0769230769231 & everything that exists anywhere\\ \hline
\multicolumn{4}{|c|}{Lenin: Imperialism, the Highes - lch similarity} \\ \hline
capitalism.n.01 & capitalism.n.01 & 3.63758615973 & an economic system based on private ownership of capital\\ \hline
proletarian.n.01 & great.n.01 & 2.02814824729 & a person who has achieved distinction and honor in some field\\ \hline
hair.n.01 & universe.n.01 & 2.02814824729 & everything that exists anywhere\\ \hline
money.n.01 & penny.n.02 & 1.8458266905 & a coin worth one-hundredth of the value of the basic unit\\ \hline
practice.n.03 & monopoly.n.01 & 1.8458266905 & (economics) a market in which there are many buyers but only one seller\\ \hline
toilet.n.04 & monopoly.n.01 & 1.69167601067 & (economics) a market in which there are many buyers but only one seller\\ \hline
war.n.01 & monopoly.n.01 & 1.69167601067 & (economics) a market in which there are many buyers but only one seller\\ \hline
revolution.n.01 & monopoly.n.01 & 1.55814461805 & (economics) a market in which there are many buyers but only one seller\\ \hline
theory.n.01 & imperialism.n.01 & 1.55814461805 & a policy of extending your rule over foreign countries\\ \hline
television receiver.n.01 & product.n.02 & 1.44036158239 & an artifact that has been created by someone or some process\\ \hline
car.n.01 & product.n.02 & 1.44036158239 & an artifact that has been created by someone or some process\\ \hline
philosophy.n.02 & policy.n.01 & 1.44036158239 & a plan of action adopted by an individual or social group\\ \hline
materialism.n.02 & policy.n.01 & 1.44036158239 & a plan of action adopted by an individual or social group\\ \hline
strategy.n.02 & policy.n.01 & 1.44036158239 & a plan of action adopted by an individual or social group\\ \hline
idealism.n.01 & policy.n.01 & 1.44036158239 & a plan of action adopted by an individual or social group\\ \hline
imperialism.n.02 & policy.n.01 & 1.33500106673 & a plan of action adopted by an individual or social group\\ \hline
lion.n.01 & great.n.01 & 1.07263680226 & a person who has achieved distinction and honor in some field\\ \hline
sugar.n.01 & universe.n.01 & 1.07263680226 & everything that exists anywhere\\ \hline
\multicolumn{4}{|c|}{Lenin: Imperialism, the Highes - res similarity} \\ \hline
capitalism.n.01 & capitalism.n.01 & 11.6930111908 & an economic system based on private ownership of capital\\ \hline
money.n.01 & penny.n.02 & 6.2549318819 & a coin worth one-hundredth of the value of the basic unit\\ \hline
theory.n.01 & imperialism.n.01 & 5.678990458 & a policy of extending your rule over foreign countries\\ \hline
hair.n.01 & universe.n.01 & 4.94037581121 & everything that exists anywhere\\ \hline
philosophy.n.02 & policy.n.01 & 3.29228159341 & a plan of action adopted by an individual or social group\\ \hline
materialism.n.02 & policy.n.01 & 3.29228159341 & a plan of action adopted by an individual or social group\\ \hline
strategy.n.02 & policy.n.01 & 3.29228159341 & a plan of action adopted by an individual or social group\\ \hline
idealism.n.01 & policy.n.01 & 3.29228159341 & a plan of action adopted by an individual or social group\\ \hline
toilet.n.04 & monopoly.n.01 & 3.14352023311 & (economics) a market in which there are many buyers but only one seller\\ \hline
imperialism.n.02 & imperialism.n.01 & 2.64452085446 & a policy of extending your rule over foreign countries\\ \hline
proletarian.n.01 & great.n.01 & 2.33354524374 & a person who has achieved distinction and honor in some field\\ \hline
war.n.01 & monopoly.n.01 & 2.33223775508 & (economics) a market in which there are many buyers but only one seller\\ \hline
practice.n.03 & monopoly.n.01 & 2.33223775508 & (economics) a market in which there are many buyers but only one seller\\ \hline
television receiver.n.01 & product.n.02 & 2.30584873145 & an artifact that has been created by someone or some process\\ \hline
car.n.01 & product.n.02 & 2.30584873145 & an artifact that has been created by someone or some process\\ \hline
lion.n.01 & great.n.01 & 2.22415047123 & a person who has achieved distinction and honor in some field\\ \hline
revolution.n.01 & monopoly.n.01 & 2.03632771779 & (economics) a market in which there are many buyers but only one seller\\ \hline
sugar.n.01 & universe.n.01 & 0.801759114954 & everything that exists anywhere\\ \hline
\multicolumn{4}{|c|}{Lenin: Imperialism, the Highes - jcn similarity} \\ \hline
capitalism.n.01 & capitalism.n.01 & 1e+300 & an economic system based on private ownership of capital\\ \hline
money.n.01 & penny.n.02 & 0.220543285308 & a coin worth one-hundredth of the value of the basic unit\\ \hline
hair.n.01 & universe.n.01 & 0.185089245333 & everything that exists anywhere\\ \hline
car.n.01 & product.n.02 & 0.118562212358 & an artifact that has been created by someone or some process\\ \hline
philosophy.n.02 & policy.n.01 & 0.0941972384174 & a plan of action adopted by an individual or social group\\ \hline
television receiver.n.01 & product.n.02 & 0.0922281130906 & an artifact that has been created by someone or some process\\ \hline
theory.n.01 & imperialism.n.01 & 0.0849319922588 & a policy of extending your rule over foreign countries\\ \hline
proletarian.n.01 & product.n.02 & 0.0827408100653 & an artifact that has been created by someone or some process\\ \hline
war.n.01 & monopoly.n.01 & 0.0769380100151 & (economics) a market in which there are many buyers but only one seller\\ \hline
lion.n.01 & product.n.02 & 0.0745010379336 & an artifact that has been created by someone or some process\\ \hline
strategy.n.02 & policy.n.01 & 0.0729992753311 & a plan of action adopted by an individual or social group\\ \hline
practice.n.03 & monopoly.n.01 & 0.0696858132394 & (economics) a market in which there are many buyers but only one seller\\ \hline
toilet.n.04 & monopoly.n.01 & 0.0675356263653 & (economics) a market in which there are many buyers but only one seller\\ \hline
materialism.n.02 & policy.n.01 & 0.0673677474158 & a plan of action adopted by an individual or social group\\ \hline
sugar.n.01 & product.n.02 & 0.0665737679313 & an artifact that has been created by someone or some process\\ \hline
revolution.n.01 & one.n.01 & 0.0654799767733 & the smallest whole number or a numeral representing this number\\ \hline
idealism.n.01 & policy.n.01 & 0.063730003856 & a plan of action adopted by an individual or social group\\ \hline
imperialism.n.02 & policy.n.01 & 0.0574970955164 & a plan of action adopted by an individual or social group\\ \hline
\multicolumn{4}{|c|}{Lenin: Imperialism, the Highes - lin similarity} \\ \hline
capitalism.n.01 & capitalism.n.01 & 1.0 & an economic system based on private ownership of capital\\ \hline
money.n.01 & penny.n.02 & 0.733969426851 & a coin worth one-hundredth of the value of the basic unit\\ \hline
hair.n.01 & universe.n.01 & 0.646495819603 & everything that exists anywhere\\ \hline
theory.n.01 & imperialism.n.01 & 0.491005027416 & a policy of extending your rule over foreign countries\\ \hline
philosophy.n.02 & policy.n.01 & 0.382810406388 & a plan of action adopted by an individual or social group\\ \hline
car.n.01 & product.n.02 & 0.353492745777 & an artifact that has been created by someone or some process\\ \hline
strategy.n.02 & policy.n.01 & 0.324629309412 & a plan of action adopted by an individual or social group\\ \hline
materialism.n.02 & policy.n.01 & 0.307281190083 & a plan of action adopted by an individual or social group\\ \hline
television receiver.n.01 & product.n.02 & 0.298407179857 & an artifact that has been created by someone or some process\\ \hline
toilet.n.04 & monopoly.n.01 & 0.298048188669 & (economics) a market in which there are many buyers but only one seller\\ \hline
idealism.n.01 & policy.n.01 & 0.295593207225 & a plan of action adopted by an individual or social group\\ \hline
war.n.01 & monopoly.n.01 & 0.264097390196 & (economics) a market in which there are many buyers but only one seller\\ \hline
proletarian.n.01 & great.n.01 & 0.262634686841 & a person who has achieved distinction and honor in some field\\ \hline
practice.n.03 & monopoly.n.01 & 0.245310227149 & (economics) a market in which there are many buyers but only one seller\\ \hline
imperialism.n.02 & policy.n.01 & 0.233190306342 & a plan of action adopted by an individual or social group\\ \hline
lion.n.01 & great.n.01 & 0.232810355981 & a person who has achieved distinction and honor in some field\\ \hline
revolution.n.01 & monopoly.n.01 & 0.205174493173 & (economics) a market in which there are many buyers but only one seller\\ \hline
sugar.n.01 & product.n.02 & 0.0964554176079 & an artifact that has been created by someone or some process\\ \hline
\end{tabular}
\end{center}