### 2014-03-01 paper/analysis_lenin_materialism_and_empirio-.tex

    \begin{center}
      \begin{tabular}{ | l | l | l | l |}
        \hline
        Choosen Sensens & Closest Top Sense & similarity & Top Sense Gloss \\ \hline
\multicolumn{4}{|c|}{Lenin: MATERIALISM and EMPIRIO - path similarity} \\ \hline
revolution.n.01 & experience.n.03 & 0.25 & an event as apprehended\\ \hline
proletarian.n.01 & matter.n.03 & 0.166666666667 & that which has mass and occupies space\\ \hline
philosophy.n.02 & physics.n.01 & 0.166666666667 & the science of matter and energy and their interactions\\ \hline
war.n.01 & experience.n.03 & 0.166666666667 & an event as apprehended\\ \hline
strategy.n.02 & physics.n.01 & 0.166666666667 & the science of matter and energy and their interactions\\ \hline
practice.n.03 & experience.n.03 & 0.166666666667 & an event as apprehended\\ \hline
money.n.01 & beginning.n.02 & 0.142857142857 & the time at which something is supposed to begin\\ \hline
materialism.n.02 & aim.n.02 & 0.142857142857 & the goal intended to be attained (and which is believed to be attainable)\\ \hline
idealism.n.01 & aim.n.02 & 0.142857142857 & the goal intended to be attained (and which is believed to be attainable)\\ \hline
imperialism.n.02 & aim.n.02 & 0.125 & the goal intended to be attained (and which is believed to be attainable)\\ \hline
sugar.n.01 & matter.n.03 & 0.125 & that which has mass and occupies space\\ \hline
theory.n.01 & sensation.n.01 & 0.125 & an unelaborated elementary awareness of stimulation\\ \hline
hair.n.01 & space.n.02 & 0.125 & an empty area (usually bounded in some way between things)\\ \hline
capitalism.n.01 & time.n.05 & 0.125 & the continuum of experience in which events pass from the future through the present to the past\\ \hline
toilet.n.04 & experience.n.03 & 0.111111111111 & an event as apprehended\\ \hline
television receiver.n.01 & space.n.02 & 0.0909090909091 & an empty area (usually bounded in some way between things)\\ \hline
car.n.01 & space.n.02 & 0.0909090909091 & an empty area (usually bounded in some way between things)\\ \hline
lion.n.01 & kant.n.01 & 0.0666666666667 & influential German idealist philosopher (1724-1804)\\ \hline
\multicolumn{4}{|c|}{Lenin: MATERIALISM and EMPIRIO - lch similarity} \\ \hline
revolution.n.01 & experience.n.03 & 2.25129179861 & an event as apprehended\\ \hline
proletarian.n.01 & matter.n.03 & 1.8458266905 & that which has mass and occupies space\\ \hline
philosophy.n.02 & physics.n.01 & 1.8458266905 & the science of matter and energy and their interactions\\ \hline
war.n.01 & experience.n.03 & 1.8458266905 & an event as apprehended\\ \hline
strategy.n.02 & physics.n.01 & 1.8458266905 & the science of matter and energy and their interactions\\ \hline
practice.n.03 & experience.n.03 & 1.8458266905 & an event as apprehended\\ \hline
money.n.01 & beginning.n.02 & 1.69167601067 & the time at which something is supposed to begin\\ \hline
materialism.n.02 & aim.n.02 & 1.69167601067 & the goal intended to be attained (and which is believed to be attainable)\\ \hline
idealism.n.01 & aim.n.02 & 1.69167601067 & the goal intended to be attained (and which is believed to be attainable)\\ \hline
imperialism.n.02 & aim.n.02 & 1.55814461805 & the goal intended to be attained (and which is believed to be attainable)\\ \hline
sugar.n.01 & matter.n.03 & 1.55814461805 & that which has mass and occupies space\\ \hline
theory.n.01 & sensation.n.01 & 1.55814461805 & an unelaborated elementary awareness of stimulation\\ \hline
hair.n.01 & space.n.02 & 1.55814461805 & an empty area (usually bounded in some way between things)\\ \hline
capitalism.n.01 & time.n.05 & 1.55814461805 & the continuum of experience in which events pass from the future through the present to the past\\ \hline
toilet.n.04 & experience.n.03 & 1.44036158239 & an event as apprehended\\ \hline
television receiver.n.01 & space.n.02 & 1.23969088693 & an empty area (usually bounded in some way between things)\\ \hline
car.n.01 & space.n.02 & 1.23969088693 & an empty area (usually bounded in some way between things)\\ \hline
lion.n.01 & kant.n.01 & 0.929535958624 & influential German idealist philosopher (1724-1804)\\ \hline
\multicolumn{4}{|c|}{Lenin: MATERIALISM and EMPIRIO - res similarity} \\ \hline
philosophy.n.02 & physics.n.01 & 5.58877614988 & the science of matter and energy and their interactions\\ \hline
strategy.n.02 & physics.n.01 & 5.58877614988 & the science of matter and energy and their interactions\\ \hline
theory.n.01 & sensation.n.01 & 4.38484522248 & an unelaborated elementary awareness of stimulation\\ \hline
revolution.n.01 & experience.n.03 & 4.05348916032 & an event as apprehended\\ \hline
sugar.n.01 & matter.n.03 & 3.3367059068 & that which has mass and occupies space\\ \hline
materialism.n.02 & physics.n.01 & 3.29228159341 & the science of matter and energy and their interactions\\ \hline
idealism.n.01 & physics.n.01 & 3.29228159341 & the science of matter and energy and their interactions\\ \hline
capitalism.n.01 & sociable.n.01 & 2.85529368615 & a party of people assembled to promote sociability and communal activity\\ \hline
money.n.01 & beginning.n.02 & 2.69110280536 & the time at which something is supposed to begin\\ \hline
imperialism.n.02 & physics.n.01 & 2.64452085446 & the science of matter and energy and their interactions\\ \hline
proletarian.n.01 & kant.n.01 & 2.33354524374 & influential German idealist philosopher (1724-1804)\\ \hline
toilet.n.04 & motion.n.03 & 2.33223775508 & a change of position that does not entail a change of location\\ \hline
war.n.01 & motion.n.03 & 2.33223775508 & a change of position that does not entail a change of location\\ \hline
practice.n.03 & motion.n.03 & 2.33223775508 & a change of position that does not entail a change of location\\ \hline
television receiver.n.01 & book.n.01 & 2.30584873145 & a written work or composition that has been published (printed on pages bound together)\\ \hline
car.n.01 & book.n.01 & 2.30584873145 & a written work or composition that has been published (printed on pages bound together)\\ \hline
lion.n.01 & kant.n.01 & 2.22415047123 & influential German idealist philosopher (1724-1804)\\ \hline
hair.n.01 & kant.n.01 & 1.53183374322 & influential German idealist philosopher (1724-1804)\\ \hline
\multicolumn{4}{|c|}{Lenin: MATERIALISM and EMPIRIO - jcn similarity} \\ \hline
philosophy.n.02 & physics.n.01 & 0.145932703696 & the science of matter and energy and their interactions\\ \hline
theory.n.01 & sensation.n.01 & 0.128286683846 & an unelaborated elementary awareness of stimulation\\ \hline
sugar.n.01 & matter.n.03 & 0.127664075736 & that which has mass and occupies space\\ \hline
revolution.n.01 & experience.n.03 & 0.111569245855 & an event as apprehended\\ \hline
money.n.01 & moment.n.01 & 0.107804699133 & a particular point in time\\ \hline
car.n.01 & matter.n.03 & 0.107243331568 & that which has mass and occupies space\\ \hline
hair.n.01 & matter.n.03 & 0.104557075904 & that which has mass and occupies space\\ \hline
strategy.n.02 & physics.n.01 & 0.100652051627 & the science of matter and energy and their interactions\\ \hline
war.n.01 & motion.n.03 & 0.100254651557 & a change of position that does not entail a change of location\\ \hline
practice.n.03 & motion.n.03 & 0.0882827134582 & a change of position that does not entail a change of location\\ \hline
proletarian.n.01 & matter.n.03 & 0.0875035868175 & that which has mass and occupies space\\ \hline
television receiver.n.01 & matter.n.03 & 0.085230569567 & that which has mass and occupies space\\ \hline
lion.n.01 & matter.n.03 & 0.0783404338338 & that which has mass and occupies space\\ \hline
materialism.n.02 & aim.n.02 & 0.0778269817906 & the goal intended to be attained (and which is believed to be attainable)\\ \hline
toilet.n.04 & motion.n.03 & 0.0745896410117 & a change of position that does not entail a change of location\\ \hline
idealism.n.01 & aim.n.02 & 0.0730123547216 & the goal intended to be attained (and which is believed to be attainable)\\ \hline
capitalism.n.01 & sociable.n.01 & 0.0678992095349 & a party of people assembled to promote sociability and communal activity\\ \hline
imperialism.n.02 & aim.n.02 & 0.0649464570442 & the goal intended to be attained (and which is believed to be attainable)\\ \hline
\multicolumn{4}{|c|}{Lenin: MATERIALISM and EMPIRIO - lin similarity} \\ \hline
philosophy.n.02 & physics.n.01 & 0.619941000616 & the science of matter and energy and their interactions\\ \hline
strategy.n.02 & physics.n.01 & 0.529421413479 & the science of matter and energy and their interactions\\ \hline
theory.n.01 & sensation.n.01 & 0.529419405948 & an unelaborated elementary awareness of stimulation\\ \hline
revolution.n.01 & experience.n.03 & 0.474924895953 & an event as apprehended\\ \hline
sugar.n.01 & matter.n.03 & 0.460030062094 & that which has mass and occupies space\\ \hline
money.n.01 & moment.n.01 & 0.367179548114 & a particular point in time\\ \hline
materialism.n.02 & aim.n.02 & 0.33882403792 & the goal intended to be attained (and which is believed to be attainable)\\ \hline
idealism.n.01 & aim.n.02 & 0.324668589607 & the goal intended to be attained (and which is believed to be attainable)\\ \hline
war.n.01 & motion.n.03 & 0.318631846502 & a change of position that does not entail a change of location\\ \hline
car.n.01 & book.n.01 & 0.308668677605 & a written work or composition that has been published (printed on pages bound together)\\ \hline
practice.n.03 & motion.n.03 & 0.291680639247 & a change of position that does not entail a change of location\\ \hline
capitalism.n.01 & sociable.n.01 & 0.27940619133 & a party of people assembled to promote sociability and communal activity\\ \hline
television receiver.n.01 & book.n.01 & 0.265820778874 & a written work or composition that has been published (printed on pages bound together)\\ \hline
toilet.n.04 & motion.n.03 & 0.258117063881 & a change of position that does not entail a change of location\\ \hline
imperialism.n.02 & aim.n.02 & 0.255677978749 & the goal intended to be attained (and which is believed to be attainable)\\ \hline
hair.n.01 & book.n.01 & 0.20182033649 & a written work or composition that has been published (printed on pages bound together)\\ \hline
proletarian.n.01 & kant.n.01 & 0.198875753057 & influential German idealist philosopher (1724-1804)\\ \hline
lion.n.01 & kant.n.01 & 0.179337574844 & influential German idealist philosopher (1724-1804)\\ \hline
\end{tabular}
\end{center}