### 2014-03-01 paper/analysis_lenin_what_is_to_be_done?.tex

    \begin{center}
      \begin{tabular}{ | l | l | l | l |}
        \hline
        Choosen Sensens & Closest Top Sense & similarity & Top Sense Gloss \\ \hline
\multicolumn{4}{|c|}{Lenin: What Is To Be Done? - path similarity} \\ \hline
imperialism.n.02 & political orientation.n.01 & 0.5 & an orientation that characterizes the thinking of a group or nation\\ \hline
proletarian.n.01 & worker.n.01 & 0.25 & a person who works at a specific occupation\\ \hline
war.n.01 & conflict.n.01 & 0.25 & an open clash between two opposing groups (or individuals)\\ \hline
practice.n.03 & conflict.n.01 & 0.2 & an open clash between two opposing groups (or individuals)\\ \hline
revolution.n.01 & conflict.n.01 & 0.166666666667 & an open clash between two opposing groups (or individuals)\\ \hline
toilet.n.04 & trade.n.02 & 0.142857142857 & the skilled practice of a practical occupation\\ \hline
money.n.01 & mass.n.01 & 0.142857142857 & the property of a body that causes it to have weight in a gravitational field\\ \hline
capitalism.n.01 & organization.n.01 & 0.142857142857 & a group of people who work together\\ \hline
hair.n.01 & working.n.01 & 0.125 & a mine or quarry that is being or has been worked\\ \hline
television receiver.n.01 & working.n.01 & 0.111111111111 & a mine or quarry that is being or has been worked\\ \hline
car.n.01 & working.n.01 & 0.111111111111 & a mine or quarry that is being or has been worked\\ \hline
philosophy.n.02 & political orientation.n.01 & 0.111111111111 & an orientation that characterizes the thinking of a group or nation\\ \hline
materialism.n.02 & political orientation.n.01 & 0.111111111111 & an orientation that characterizes the thinking of a group or nation\\ \hline
strategy.n.02 & political orientation.n.01 & 0.111111111111 & an orientation that characterizes the thinking of a group or nation\\ \hline
idealism.n.01 & political orientation.n.01 & 0.111111111111 & an orientation that characterizes the thinking of a group or nation\\ \hline
theory.n.01 & political orientation.n.01 & 0.111111111111 & an orientation that characterizes the thinking of a group or nation\\ \hline
sugar.n.01 & local anesthetic.n.01 & 0.0909090909091 & anesthetic that numbs a particular area of the body\\ \hline
lion.n.01 & worker.n.01 & 0.0833333333333 & a person who works at a specific occupation\\ \hline
\multicolumn{4}{|c|}{Lenin: What Is To Be Done? - lch similarity} \\ \hline
imperialism.n.02 & political orientation.n.01 & 2.94443897917 & an orientation that characterizes the thinking of a group or nation\\ \hline
proletarian.n.01 & worker.n.01 & 2.25129179861 & a person who works at a specific occupation\\ \hline
war.n.01 & conflict.n.01 & 2.25129179861 & an open clash between two opposing groups (or individuals)\\ \hline
practice.n.03 & conflict.n.01 & 2.02814824729 & an open clash between two opposing groups (or individuals)\\ \hline
revolution.n.01 & conflict.n.01 & 1.8458266905 & an open clash between two opposing groups (or individuals)\\ \hline
toilet.n.04 & trade.n.02 & 1.69167601067 & the skilled practice of a practical occupation\\ \hline
money.n.01 & mass.n.01 & 1.69167601067 & the property of a body that causes it to have weight in a gravitational field\\ \hline
capitalism.n.01 & organization.n.01 & 1.69167601067 & a group of people who work together\\ \hline
hair.n.01 & working.n.01 & 1.55814461805 & a mine or quarry that is being or has been worked\\ \hline
television receiver.n.01 & working.n.01 & 1.44036158239 & a mine or quarry that is being or has been worked\\ \hline
car.n.01 & working.n.01 & 1.44036158239 & a mine or quarry that is being or has been worked\\ \hline
philosophy.n.02 & political orientation.n.01 & 1.44036158239 & an orientation that characterizes the thinking of a group or nation\\ \hline
materialism.n.02 & political orientation.n.01 & 1.44036158239 & an orientation that characterizes the thinking of a group or nation\\ \hline
strategy.n.02 & political orientation.n.01 & 1.44036158239 & an orientation that characterizes the thinking of a group or nation\\ \hline
idealism.n.01 & political orientation.n.01 & 1.44036158239 & an orientation that characterizes the thinking of a group or nation\\ \hline
theory.n.01 & political orientation.n.01 & 1.44036158239 & an orientation that characterizes the thinking of a group or nation\\ \hline
sugar.n.01 & local anesthetic.n.01 & 1.23969088693 & anesthetic that numbs a particular area of the body\\ \hline
lion.n.01 & worker.n.01 & 1.15267950994 & a person who works at a specific occupation\\ \hline
\multicolumn{4}{|c|}{Lenin: What Is To Be Done? - res similarity} \\ \hline
imperialism.n.02 & political orientation.n.01 & 8.60196873746 & an orientation that characterizes the thinking of a group or nation\\ \hline
sugar.n.01 & local anesthetic.n.01 & 4.61519673162 & anesthetic that numbs a particular area of the body\\ \hline
war.n.01 & conflict.n.01 & 4.52385364841 & an open clash between two opposing groups (or individuals)\\ \hline
television receiver.n.01 & newspaper.n.01 & 3.25767902589 & a daily or weekly publication on folded sheets; contains news and articles and advertisements\\ \hline
car.n.01 & newspaper.n.01 & 3.25767902589 & a daily or weekly publication on folded sheets; contains news and articles and advertisements\\ \hline
toilet.n.04 & trade.n.02 & 3.14352023311 & the skilled practice of a practical occupation\\ \hline
capitalism.n.01 & organization.n.01 & 2.85529368615 & a group of people who work together\\ \hline
money.n.01 & mass.n.01 & 2.69110280536 & the property of a body that causes it to have weight in a gravitational field\\ \hline
philosophy.n.02 & political orientation.n.01 & 2.64452085446 & an orientation that characterizes the thinking of a group or nation\\ \hline
materialism.n.02 & political orientation.n.01 & 2.64452085446 & an orientation that characterizes the thinking of a group or nation\\ \hline
strategy.n.02 & political orientation.n.01 & 2.64452085446 & an orientation that characterizes the thinking of a group or nation\\ \hline
idealism.n.01 & political orientation.n.01 & 2.64452085446 & an orientation that characterizes the thinking of a group or nation\\ \hline
theory.n.01 & political orientation.n.01 & 2.64452085446 & an orientation that characterizes the thinking of a group or nation\\ \hline
proletarian.n.01 & worker.n.01 & 2.33354524374 & a person who works at a specific occupation\\ \hline
practice.n.03 & trade.n.02 & 2.33223775508 & the skilled practice of a practical occupation\\ \hline
lion.n.01 & worker.n.01 & 2.22415047123 & a person who works at a specific occupation\\ \hline
revolution.n.01 & trade.n.02 & 2.03632771779 & the skilled practice of a practical occupation\\ \hline
hair.n.01 & worker.n.01 & 1.53183374322 & a person who works at a specific occupation\\ \hline
\multicolumn{4}{|c|}{Lenin: What Is To Be Done? - jcn similarity} \\ \hline
imperialism.n.02 & political orientation.n.01 & 0.193405491818 & an orientation that characterizes the thinking of a group or nation\\ \hline
war.n.01 & conflict.n.01 & 0.152829762599 & an open clash between two opposing groups (or individuals)\\ \hline
car.n.01 & worker.n.01 & 0.110550521246 & a person who works at a specific occupation\\ \hline
hair.n.01 & worker.n.01 & 0.107698235698 & a person who works at a specific occupation\\ \hline
proletarian.n.01 & worker.n.01 & 0.104758893715 & a person who works at a specific occupation\\ \hline
capitalism.n.01 & organization.n.01 & 0.101121795595 & a group of people who work together\\ \hline
sugar.n.01 & local anesthetic.n.01 & 0.0929714058017 & anesthetic that numbs a particular area of the body\\ \hline
money.n.01 & organization.n.01 & 0.0929667353193 & a group of people who work together\\ \hline
philosophy.n.02 & organization.n.01 & 0.0908595602215 & a group of people who work together\\ \hline
lion.n.01 & worker.n.01 & 0.0900802228812 & a person who works at a specific occupation\\ \hline
practice.n.03 & motion.n.03 & 0.0882827134582 & a change of position that does not entail a change of location\\ \hline
television receiver.n.01 & worker.n.01 & 0.0873062903722 & a person who works at a specific occupation\\ \hline
theory.n.01 & organization.n.01 & 0.0828237113739 & a group of people who work together\\ \hline
revolution.n.01 & motion.n.03 & 0.0784056470694 & a change of position that does not entail a change of location\\ \hline
toilet.n.04 & motion.n.03 & 0.0745896410117 & a change of position that does not entail a change of location\\ \hline
strategy.n.02 & organization.n.01 & 0.0709786674118 & a group of people who work together\\ \hline
materialism.n.02 & organization.n.01 & 0.0656431911358 & a group of people who work together\\ \hline
idealism.n.01 & organization.n.01 & 0.0621845289629 & a group of people who work together\\ \hline
\multicolumn{4}{|c|}{Lenin: What Is To Be Done? - lin similarity} \\ \hline
imperialism.n.02 & political orientation.n.01 & 0.768910941363 & an orientation that characterizes the thinking of a group or nation\\ \hline
war.n.01 & conflict.n.01 & 0.580318438615 & an open clash between two opposing groups (or individuals)\\ \hline
sugar.n.01 & local anesthetic.n.01 & 0.461833980698 & anesthetic that numbs a particular area of the body\\ \hline
car.n.01 & newspaper.n.01 & 0.411156952767 & a daily or weekly publication on folded sheets; contains news and articles and advertisements\\ \hline
capitalism.n.01 & organization.n.01 & 0.366071452787 & a group of people who work together\\ \hline
television receiver.n.01 & newspaper.n.01 & 0.356914166707 & a daily or weekly publication on folded sheets; contains news and articles and advertisements\\ \hline
money.n.01 & mass.n.01 & 0.331362497131 & the property of a body that causes it to have weight in a gravitational field\\ \hline
proletarian.n.01 & worker.n.01 & 0.328371898491 & a person who works at a specific occupation\\ \hline
philosophy.n.02 & political orientation.n.01 & 0.31307622756 & an orientation that characterizes the thinking of a group or nation\\ \hline
toilet.n.04 & trade.n.02 & 0.312879932381 & the skilled practice of a practical occupation\\ \hline
theory.n.01 & political orientation.n.01 & 0.294463460062 & an orientation that characterizes the thinking of a group or nation\\ \hline
practice.n.03 & motion.n.03 & 0.291680639247 & a change of position that does not entail a change of location\\ \hline
lion.n.01 & worker.n.01 & 0.286073258452 & a person who works at a specific occupation\\ \hline
strategy.n.02 & political orientation.n.01 & 0.264762834975 & an orientation that characterizes the thinking of a group or nation\\ \hline
materialism.n.02 & political orientation.n.01 & 0.25040844726 & an orientation that characterizes the thinking of a group or nation\\ \hline
hair.n.01 & worker.n.01 & 0.248092928538 & a person who works at a specific occupation\\ \hline
revolution.n.01 & motion.n.03 & 0.242033306584 & a change of position that does not entail a change of location\\ \hline
idealism.n.01 & political orientation.n.01 & 0.240750703939 & an orientation that characterizes the thinking of a group or nation\\ \hline
\end{tabular}
\end{center}