### 2014-03-01 paper/analysis_the_state_and_revolution.tex

    \begin{center}
      \begin{tabular}{ | l | l | l | l |}
        \hline
        Choosen Sensens & Closest Top Sense & similarity & Top Sense Gloss \\ \hline
\multicolumn{4}{|c|}{The State and Revolution - path similarity} \\ \hline
revolution.n.01 & revolution.n.01 & 1.0 & a drastic and far-reaching change in ways of thinking and behaving\\ \hline
imperialism.n.02 & communism.n.02 & 0.333333333333 & a political theory favoring collectivism in a classless society\\ \hline
proletarian.n.01 & peer.n.01 & 0.25 & a person who is of equal standing with another in a group\\ \hline
practice.n.03 & work.n.01 & 0.2 & activity directed toward making or doing something\\ \hline
toilet.n.04 & work.n.01 & 0.166666666667 & activity directed toward making or doing something\\ \hline
war.n.01 & work.n.01 & 0.166666666667 & activity directed toward making or doing something\\ \hline
money.n.01 & phase.n.01 & 0.125 & any distinct time period in a sequence of events\\ \hline
philosophy.n.02 & law.n.04 & 0.125 & a generalization that describes recurring facts or events in nature\\ \hline
materialism.n.02 & law.n.04 & 0.125 & a generalization that describes recurring facts or events in nature\\ \hline
strategy.n.02 & law.n.04 & 0.125 & a generalization that describes recurring facts or events in nature\\ \hline
idealism.n.01 & law.n.04 & 0.125 & a generalization that describes recurring facts or events in nature\\ \hline
capitalism.n.01 & democracy.n.02 & 0.125 & a political system in which the supreme power lies in a body of citizens who can elect people to represent them\\ \hline
hair.n.01 & peer.n.01 & 0.111111111111 & a person who is of equal standing with another in a group\\ \hline
theory.n.01 & experience.n.03 & 0.1 & an event as apprehended\\ \hline
television receiver.n.01 & peer.n.01 & 0.0833333333333 & a person who is of equal standing with another in a group\\ \hline
car.n.01 & peer.n.01 & 0.0833333333333 & a person who is of equal standing with another in a group\\ \hline
lion.n.01 & peer.n.01 & 0.0833333333333 & a person who is of equal standing with another in a group\\ \hline
sugar.n.01 & peer.n.01 & 0.0833333333333 & a person who is of equal standing with another in a group\\ \hline
\multicolumn{4}{|c|}{The State and Revolution - lch similarity} \\ \hline
revolution.n.01 & revolution.n.01 & 3.63758615973 & a drastic and far-reaching change in ways of thinking and behaving\\ \hline
imperialism.n.02 & communism.n.02 & 2.53897387106 & a political theory favoring collectivism in a classless society\\ \hline
proletarian.n.01 & peer.n.01 & 2.25129179861 & a person who is of equal standing with another in a group\\ \hline
practice.n.03 & work.n.01 & 2.02814824729 & activity directed toward making or doing something\\ \hline
toilet.n.04 & work.n.01 & 1.8458266905 & activity directed toward making or doing something\\ \hline
war.n.01 & work.n.01 & 1.8458266905 & activity directed toward making or doing something\\ \hline
money.n.01 & phase.n.01 & 1.55814461805 & any distinct time period in a sequence of events\\ \hline
philosophy.n.02 & law.n.04 & 1.55814461805 & a generalization that describes recurring facts or events in nature\\ \hline
materialism.n.02 & law.n.04 & 1.55814461805 & a generalization that describes recurring facts or events in nature\\ \hline
strategy.n.02 & law.n.04 & 1.55814461805 & a generalization that describes recurring facts or events in nature\\ \hline
idealism.n.01 & law.n.04 & 1.55814461805 & a generalization that describes recurring facts or events in nature\\ \hline
capitalism.n.01 & democracy.n.02 & 1.55814461805 & a political system in which the supreme power lies in a body of citizens who can elect people to represent them\\ \hline
hair.n.01 & peer.n.01 & 1.44036158239 & a person who is of equal standing with another in a group\\ \hline
theory.n.01 & experience.n.03 & 1.33500106673 & an event as apprehended\\ \hline
television receiver.n.01 & peer.n.01 & 1.15267950994 & a person who is of equal standing with another in a group\\ \hline
car.n.01 & peer.n.01 & 1.15267950994 & a person who is of equal standing with another in a group\\ \hline
lion.n.01 & peer.n.01 & 1.15267950994 & a person who is of equal standing with another in a group\\ \hline
sugar.n.01 & peer.n.01 & 1.15267950994 & a person who is of equal standing with another in a group\\ \hline
\multicolumn{4}{|c|}{The State and Revolution - res similarity} \\ \hline
revolution.n.01 & revolution.n.01 & 10.3880624692 & a drastic and far-reaching change in ways of thinking and behaving\\ \hline
imperialism.n.02 & communism.n.02 & 8.60196873746 & a political theory favoring collectivism in a classless society\\ \hline
philosophy.n.02 & law.n.04 & 3.29228159341 & a generalization that describes recurring facts or events in nature\\ \hline
materialism.n.02 & law.n.04 & 3.29228159341 & a generalization that describes recurring facts or events in nature\\ \hline
strategy.n.02 & law.n.04 & 3.29228159341 & a generalization that describes recurring facts or events in nature\\ \hline
idealism.n.01 & law.n.04 & 3.29228159341 & a generalization that describes recurring facts or events in nature\\ \hline
toilet.n.04 & work.n.01 & 3.14352023311 & activity directed toward making or doing something\\ \hline
capitalism.n.01 & club.n.02 & 2.85529368615 & a formal association of people with similar interests\\ \hline
money.n.01 & phase.n.01 & 2.69110280536 & any distinct time period in a sequence of events\\ \hline
theory.n.01 & law.n.04 & 2.64452085446 & a generalization that describes recurring facts or events in nature\\ \hline
proletarian.n.01 & engels.n.01 & 2.33354524374 & socialist who wrote the Communist Manifesto with Karl Marx in 1848 (1820-1895)\\ \hline
war.n.01 & work.n.01 & 2.33223775508 & activity directed toward making or doing something\\ \hline
practice.n.03 & work.n.01 & 2.33223775508 & activity directed toward making or doing something\\ \hline
lion.n.01 & engels.n.01 & 2.22415047123 & socialist who wrote the Communist Manifesto with Karl Marx in 1848 (1820-1895)\\ \hline
television receiver.n.01 & engels.n.01 & 1.53183374322 & socialist who wrote the Communist Manifesto with Karl Marx in 1848 (1820-1895)\\ \hline
car.n.01 & engels.n.01 & 1.53183374322 & socialist who wrote the Communist Manifesto with Karl Marx in 1848 (1820-1895)\\ \hline
hair.n.01 & engels.n.01 & 1.53183374322 & socialist who wrote the Communist Manifesto with Karl Marx in 1848 (1820-1895)\\ \hline
sugar.n.01 & engels.n.01 & 0.801759114954 & socialist who wrote the Communist Manifesto with Karl Marx in 1848 (1820-1895)\\ \hline
\multicolumn{4}{|c|}{The State and Revolution - jcn similarity} \\ \hline
revolution.n.01 & revolution.n.01 & 1e+300 & a drastic and far-reaching change in ways of thinking and behaving\\ \hline
imperialism.n.02 & communism.n.02 & 0.14975874591 & a political theory favoring collectivism in a classless society\\ \hline
war.n.01 & work.n.01 & 0.119881984187 & activity directed toward making or doing something\\ \hline
philosophy.n.02 & law.n.04 & 0.103411969458 & a generalization that describes recurring facts or events in nature\\ \hline
practice.n.03 & work.n.01 & 0.103154641506 & activity directed toward making or doing something\\ \hline
toilet.n.04 & work.n.01 & 0.0985118809398 & activity directed toward making or doing something\\ \hline
money.n.01 & phase.n.01 & 0.0916158054473 & any distinct time period in a sequence of events\\ \hline
theory.n.01 & work.n.01 & 0.0910758552239 & activity directed toward making or doing something\\ \hline
car.n.01 & peer.n.01 & 0.0887294617483 & a person who is of equal standing with another in a group\\ \hline
hair.n.01 & peer.n.01 & 0.0868826421077 & a person who is of equal standing with another in a group\\ \hline
proletarian.n.01 & peer.n.01 & 0.0849595683478 & a person who is of equal standing with another in a group\\ \hline
strategy.n.02 & law.n.04 & 0.07841412673 & a generalization that describes recurring facts or events in nature\\ \hline
lion.n.01 & peer.n.01 & 0.0750424463326 & a person who is of equal standing with another in a group\\ \hline
television receiver.n.01 & peer.n.01 & 0.0731074093034 & a person who is of equal standing with another in a group\\ \hline
capitalism.n.01 & club.n.02 & 0.0729022081709 & a formal association of people with similar interests\\ \hline
materialism.n.02 & law.n.04 & 0.0719531280061 & a generalization that describes recurring facts or events in nature\\ \hline
idealism.n.01 & law.n.04 & 0.0678185221973 & a generalization that describes recurring facts or events in nature\\ \hline
sugar.n.01 & work.n.01 & 0.0625959407953 & activity directed toward making or doing something\\ \hline
\multicolumn{4}{|c|}{The State and Revolution - lin similarity} \\ \hline
revolution.n.01 & revolution.n.01 & 1.0 & a drastic and far-reaching change in ways of thinking and behaving\\ \hline
imperialism.n.02 & communism.n.02 & 0.720392353354 & a political theory favoring collectivism in a classless society\\ \hline
philosophy.n.02 & law.n.04 & 0.405088626962 & a generalization that describes recurring facts or events in nature\\ \hline
toilet.n.04 & work.n.01 & 0.382467580972 & activity directed toward making or doing something\\ \hline
war.n.01 & work.n.01 & 0.35863993877 & activity directed toward making or doing something\\ \hline
strategy.n.02 & law.n.04 & 0.340509805473 & a generalization that describes recurring facts or events in nature\\ \hline
money.n.01 & phase.n.01 & 0.330250297796 & any distinct time period in a sequence of events\\ \hline
practice.n.03 & work.n.01 & 0.324854541172 & activity directed toward making or doing something\\ \hline
materialism.n.02 & law.n.04 & 0.321472637881 & a generalization that describes recurring facts or events in nature\\ \hline
idealism.n.01 & law.n.04 & 0.308702564543 & a generalization that describes recurring facts or events in nature\\ \hline
theory.n.01 & law.n.04 & 0.305328480717 & a generalization that describes recurring facts or events in nature\\ \hline
capitalism.n.01 & club.n.02 & 0.29394209418 & a formal association of people with similar interests\\ \hline
proletarian.n.01 & peer.n.01 & 0.28393127113 & a person who is of equal standing with another in a group\\ \hline
lion.n.01 & peer.n.01 & 0.250268807542 & a person who is of equal standing with another in a group\\ \hline
car.n.01 & peer.n.01 & 0.213736072979 & a person who is of equal standing with another in a group\\ \hline
hair.n.01 & peer.n.01 & 0.210222579299 & a person who is of equal standing with another in a group\\ \hline
television receiver.n.01 & peer.n.01 & 0.182991045418 & a person who is of equal standing with another in a group\\ \hline
sugar.n.01 & peer.n.01 & 0.0895207954547 & a person who is of equal standing with another in a group\\ \hline
\end{tabular}
\end{center}