### 2014-03-02 paper/analysis_lenin_imperialism_the_highest_.tex
add initial results to the paper

    \begin{center}
      \begin{tabular}{ | l | l | l |}
        \hline
        Choosen Sensens & Closest Top Sense & similarity  \\ \hline
\multicolumn{3}{|c|}{Lenin: Imperialism, the Highes - path similarity} \\ \hline
capitalism.n.01 & capitalism.n.01 & 1.0\\ \hline
proletarian.n.01 & great.n.01 & 0.2\\ \hline
hair.n.01 & universe.n.01 & 0.2\\ \hline
money.n.01 & penny.n.02 & 0.166666666667\\ \hline
practice.n.03 & monopoly.n.01 & 0.166666666667\\ \hline
toilet.n.04 & monopoly.n.01 & 0.142857142857\\ \hline
war.n.01 & monopoly.n.01 & 0.142857142857\\ \hline
revolution.n.01 & monopoly.n.01 & 0.125\\ \hline
theory.n.01 & imperialism.n.01 & 0.125\\ \hline
television receiver.n.01 & product.n.02 & 0.111111111111\\ \hline
car.n.01 & product.n.02 & 0.111111111111\\ \hline
philosophy.n.02 & policy.n.01 & 0.111111111111\\ \hline
materialism.n.02 & policy.n.01 & 0.111111111111\\ \hline
strategy.n.02 & policy.n.01 & 0.111111111111\\ \hline
idealism.n.01 & policy.n.01 & 0.111111111111\\ \hline
imperialism.n.02 & policy.n.01 & 0.1\\ \hline
lion.n.01 & great.n.01 & 0.0769230769231\\ \hline
sugar.n.01 & universe.n.01 & 0.0769230769231\\ \hline
\multicolumn{3}{|c|}{Lenin: Imperialism, the Highes - lch similarity} \\ \hline
capitalism.n.01 & capitalism.n.01 & 3.63758615973\\ \hline
proletarian.n.01 & great.n.01 & 2.02814824729\\ \hline
hair.n.01 & universe.n.01 & 2.02814824729\\ \hline
money.n.01 & penny.n.02 & 1.8458266905\\ \hline
practice.n.03 & monopoly.n.01 & 1.8458266905\\ \hline
toilet.n.04 & monopoly.n.01 & 1.69167601067\\ \hline
war.n.01 & monopoly.n.01 & 1.69167601067\\ \hline
revolution.n.01 & monopoly.n.01 & 1.55814461805\\ \hline
theory.n.01 & imperialism.n.01 & 1.55814461805\\ \hline
television receiver.n.01 & product.n.02 & 1.44036158239\\ \hline
car.n.01 & product.n.02 & 1.44036158239\\ \hline
philosophy.n.02 & policy.n.01 & 1.44036158239\\ \hline
materialism.n.02 & policy.n.01 & 1.44036158239\\ \hline
strategy.n.02 & policy.n.01 & 1.44036158239\\ \hline
idealism.n.01 & policy.n.01 & 1.44036158239\\ \hline
imperialism.n.02 & policy.n.01 & 1.33500106673\\ \hline
lion.n.01 & great.n.01 & 1.07263680226\\ \hline
sugar.n.01 & universe.n.01 & 1.07263680226\\ \hline
\multicolumn{3}{|c|}{Lenin: Imperialism, the Highes - res similarity} \\ \hline
capitalism.n.01 & capitalism.n.01 & 11.6930111908\\ \hline
money.n.01 & penny.n.02 & 6.2549318819\\ \hline
theory.n.01 & imperialism.n.01 & 5.678990458\\ \hline
hair.n.01 & universe.n.01 & 4.94037581121\\ \hline
philosophy.n.02 & policy.n.01 & 3.29228159341\\ \hline
materialism.n.02 & policy.n.01 & 3.29228159341\\ \hline
strategy.n.02 & policy.n.01 & 3.29228159341\\ \hline
idealism.n.01 & policy.n.01 & 3.29228159341\\ \hline
toilet.n.04 & monopoly.n.01 & 3.14352023311\\ \hline
imperialism.n.02 & imperialism.n.01 & 2.64452085446\\ \hline
proletarian.n.01 & great.n.01 & 2.33354524374\\ \hline
war.n.01 & monopoly.n.01 & 2.33223775508\\ \hline
practice.n.03 & monopoly.n.01 & 2.33223775508\\ \hline
television receiver.n.01 & product.n.02 & 2.30584873145\\ \hline
car.n.01 & product.n.02 & 2.30584873145\\ \hline
lion.n.01 & great.n.01 & 2.22415047123\\ \hline
revolution.n.01 & monopoly.n.01 & 2.03632771779\\ \hline
sugar.n.01 & universe.n.01 & 0.801759114954\\ \hline
\multicolumn{3}{|c|}{Lenin: Imperialism, the Highes - jcn similarity} \\ \hline
capitalism.n.01 & capitalism.n.01 & 1e+300\\ \hline
money.n.01 & penny.n.02 & 0.220543285308\\ \hline
hair.n.01 & universe.n.01 & 0.185089245333\\ \hline
car.n.01 & product.n.02 & 0.118562212358\\ \hline
philosophy.n.02 & policy.n.01 & 0.0941972384174\\ \hline
television receiver.n.01 & product.n.02 & 0.0922281130906\\ \hline
theory.n.01 & imperialism.n.01 & 0.0849319922588\\ \hline
proletarian.n.01 & product.n.02 & 0.0827408100653\\ \hline
war.n.01 & monopoly.n.01 & 0.0769380100151\\ \hline
lion.n.01 & product.n.02 & 0.0745010379336\\ \hline
strategy.n.02 & policy.n.01 & 0.0729992753311\\ \hline
practice.n.03 & monopoly.n.01 & 0.0696858132394\\ \hline
toilet.n.04 & monopoly.n.01 & 0.0675356263653\\ \hline
materialism.n.02 & policy.n.01 & 0.0673677474158\\ \hline
sugar.n.01 & product.n.02 & 0.0665737679313\\ \hline
revolution.n.01 & one.n.01 & 0.0654799767733\\ \hline
idealism.n.01 & policy.n.01 & 0.063730003856\\ \hline
imperialism.n.02 & policy.n.01 & 0.0574970955164\\ \hline
\multicolumn{3}{|c|}{Lenin: Imperialism, the Highes - lin similarity} \\ \hline
capitalism.n.01 & capitalism.n.01 & 1.0\\ \hline
money.n.01 & penny.n.02 & 0.733969426851\\ \hline
hair.n.01 & universe.n.01 & 0.646495819603\\ \hline
theory.n.01 & imperialism.n.01 & 0.491005027416\\ \hline
philosophy.n.02 & policy.n.01 & 0.382810406388\\ \hline
car.n.01 & product.n.02 & 0.353492745777\\ \hline
strategy.n.02 & policy.n.01 & 0.324629309412\\ \hline
materialism.n.02 & policy.n.01 & 0.307281190083\\ \hline
television receiver.n.01 & product.n.02 & 0.298407179857\\ \hline
toilet.n.04 & monopoly.n.01 & 0.298048188669\\ \hline
idealism.n.01 & policy.n.01 & 0.295593207225\\ \hline
war.n.01 & monopoly.n.01 & 0.264097390196\\ \hline
proletarian.n.01 & great.n.01 & 0.262634686841\\ \hline
practice.n.03 & monopoly.n.01 & 0.245310227149\\ \hline
imperialism.n.02 & policy.n.01 & 0.233190306342\\ \hline
lion.n.01 & great.n.01 & 0.232810355981\\ \hline
revolution.n.01 & monopoly.n.01 & 0.205174493173\\ \hline
sugar.n.01 & product.n.02 & 0.0964554176079\\ \hline
\end{tabular}
\end{center}