### 2014-03-02 paper/analysis_lenin_materialism_and_empirio-.tex

    \begin{center}
      \begin{tabular}{ | l | l | l |}
        \hline
        Choosen Sensens & Closest Top Sense & similarity  \\ \hline
\multicolumn{3}{|c|}{Lenin: MATERIALISM and EMPIRIO - path similarity} \\ \hline
revolution.n.01 & experience.n.03 & 0.25\\ \hline
proletarian.n.01 & matter.n.03 & 0.166666666667\\ \hline
philosophy.n.02 & physics.n.01 & 0.166666666667\\ \hline
war.n.01 & experience.n.03 & 0.166666666667\\ \hline
strategy.n.02 & physics.n.01 & 0.166666666667\\ \hline
practice.n.03 & experience.n.03 & 0.166666666667\\ \hline
money.n.01 & beginning.n.02 & 0.142857142857\\ \hline
materialism.n.02 & aim.n.02 & 0.142857142857\\ \hline
idealism.n.01 & aim.n.02 & 0.142857142857\\ \hline
imperialism.n.02 & aim.n.02 & 0.125\\ \hline
sugar.n.01 & matter.n.03 & 0.125\\ \hline
theory.n.01 & sensation.n.01 & 0.125\\ \hline
hair.n.01 & space.n.02 & 0.125\\ \hline
capitalism.n.01 & time.n.05 & 0.125\\ \hline
toilet.n.04 & experience.n.03 & 0.111111111111\\ \hline
television receiver.n.01 & space.n.02 & 0.0909090909091\\ \hline
car.n.01 & space.n.02 & 0.0909090909091\\ \hline
lion.n.01 & kant.n.01 & 0.0666666666667\\ \hline
\multicolumn{3}{|c|}{Lenin: MATERIALISM and EMPIRIO - lch similarity} \\ \hline
revolution.n.01 & experience.n.03 & 2.25129179861\\ \hline
proletarian.n.01 & matter.n.03 & 1.8458266905\\ \hline
philosophy.n.02 & physics.n.01 & 1.8458266905\\ \hline
war.n.01 & experience.n.03 & 1.8458266905\\ \hline
strategy.n.02 & physics.n.01 & 1.8458266905\\ \hline
practice.n.03 & experience.n.03 & 1.8458266905\\ \hline
money.n.01 & beginning.n.02 & 1.69167601067\\ \hline
materialism.n.02 & aim.n.02 & 1.69167601067\\ \hline
idealism.n.01 & aim.n.02 & 1.69167601067\\ \hline
imperialism.n.02 & aim.n.02 & 1.55814461805\\ \hline
sugar.n.01 & matter.n.03 & 1.55814461805\\ \hline
theory.n.01 & sensation.n.01 & 1.55814461805\\ \hline
hair.n.01 & space.n.02 & 1.55814461805\\ \hline
capitalism.n.01 & time.n.05 & 1.55814461805\\ \hline
toilet.n.04 & experience.n.03 & 1.44036158239\\ \hline
television receiver.n.01 & space.n.02 & 1.23969088693\\ \hline
car.n.01 & space.n.02 & 1.23969088693\\ \hline
lion.n.01 & kant.n.01 & 0.929535958624\\ \hline
\multicolumn{3}{|c|}{Lenin: MATERIALISM and EMPIRIO - res similarity} \\ \hline
philosophy.n.02 & physics.n.01 & 5.58877614988\\ \hline
strategy.n.02 & physics.n.01 & 5.58877614988\\ \hline
theory.n.01 & sensation.n.01 & 4.38484522248\\ \hline
revolution.n.01 & experience.n.03 & 4.05348916032\\ \hline
sugar.n.01 & matter.n.03 & 3.3367059068\\ \hline
materialism.n.02 & physics.n.01 & 3.29228159341\\ \hline
idealism.n.01 & physics.n.01 & 3.29228159341\\ \hline
capitalism.n.01 & sociable.n.01 & 2.85529368615\\ \hline
money.n.01 & beginning.n.02 & 2.69110280536\\ \hline
imperialism.n.02 & physics.n.01 & 2.64452085446\\ \hline
proletarian.n.01 & kant.n.01 & 2.33354524374\\ \hline
toilet.n.04 & motion.n.03 & 2.33223775508\\ \hline
war.n.01 & motion.n.03 & 2.33223775508\\ \hline
practice.n.03 & motion.n.03 & 2.33223775508\\ \hline
television receiver.n.01 & book.n.01 & 2.30584873145\\ \hline
car.n.01 & book.n.01 & 2.30584873145\\ \hline
lion.n.01 & kant.n.01 & 2.22415047123\\ \hline
hair.n.01 & kant.n.01 & 1.53183374322\\ \hline
\multicolumn{3}{|c|}{Lenin: MATERIALISM and EMPIRIO - jcn similarity} \\ \hline
philosophy.n.02 & physics.n.01 & 0.145932703696\\ \hline
theory.n.01 & sensation.n.01 & 0.128286683846\\ \hline
sugar.n.01 & matter.n.03 & 0.127664075736\\ \hline
revolution.n.01 & experience.n.03 & 0.111569245855\\ \hline
money.n.01 & moment.n.01 & 0.107804699133\\ \hline
car.n.01 & matter.n.03 & 0.107243331568\\ \hline
hair.n.01 & matter.n.03 & 0.104557075904\\ \hline
strategy.n.02 & physics.n.01 & 0.100652051627\\ \hline
war.n.01 & motion.n.03 & 0.100254651557\\ \hline
practice.n.03 & motion.n.03 & 0.0882827134582\\ \hline
proletarian.n.01 & matter.n.03 & 0.0875035868175\\ \hline
television receiver.n.01 & matter.n.03 & 0.085230569567\\ \hline
lion.n.01 & matter.n.03 & 0.0783404338338\\ \hline
materialism.n.02 & aim.n.02 & 0.0778269817906\\ \hline
toilet.n.04 & motion.n.03 & 0.0745896410117\\ \hline
idealism.n.01 & aim.n.02 & 0.0730123547216\\ \hline
capitalism.n.01 & sociable.n.01 & 0.0678992095349\\ \hline
imperialism.n.02 & aim.n.02 & 0.0649464570442\\ \hline
\multicolumn{3}{|c|}{Lenin: MATERIALISM and EMPIRIO - lin similarity} \\ \hline
philosophy.n.02 & physics.n.01 & 0.619941000616\\ \hline
strategy.n.02 & physics.n.01 & 0.529421413479\\ \hline
theory.n.01 & sensation.n.01 & 0.529419405948\\ \hline
revolution.n.01 & experience.n.03 & 0.474924895953\\ \hline
sugar.n.01 & matter.n.03 & 0.460030062094\\ \hline
money.n.01 & moment.n.01 & 0.367179548114\\ \hline
materialism.n.02 & aim.n.02 & 0.33882403792\\ \hline
idealism.n.01 & aim.n.02 & 0.324668589607\\ \hline
war.n.01 & motion.n.03 & 0.318631846502\\ \hline
car.n.01 & book.n.01 & 0.308668677605\\ \hline
practice.n.03 & motion.n.03 & 0.291680639247\\ \hline
capitalism.n.01 & sociable.n.01 & 0.27940619133\\ \hline
television receiver.n.01 & book.n.01 & 0.265820778874\\ \hline
toilet.n.04 & motion.n.03 & 0.258117063881\\ \hline
imperialism.n.02 & aim.n.02 & 0.255677978749\\ \hline
hair.n.01 & book.n.01 & 0.20182033649\\ \hline
proletarian.n.01 & kant.n.01 & 0.198875753057\\ \hline
lion.n.01 & kant.n.01 & 0.179337574844\\ \hline
\end{tabular}
\end{center}