### 2014-03-02 paper/analysis_lenin_the_agrarian_programme_o.tex

    \begin{center}
      \begin{tabular}{ | l | l | l |}
        \hline
        Choosen Sensens & Closest Top Sense & similarity  \\ \hline
\multicolumn{3}{|c|}{Lenin: The Agrarian Programme  - path similarity} \\ \hline
revolution.n.01 & revolution.n.01 & 1.0\\ \hline
proletarian.n.01 & peasant.n.01 & 0.25\\ \hline
money.n.01 & bill.n.03 & 0.2\\ \hline
war.n.01 & nationalization.n.03 & 0.2\\ \hline
practice.n.03 & nationalization.n.03 & 0.166666666667\\ \hline
capitalism.n.01 & people.n.01 & 0.166666666667\\ \hline
hair.n.01 & land.n.04 & 0.142857142857\\ \hline
toilet.n.04 & nationalization.n.03 & 0.111111111111\\ \hline
philosophy.n.02 & absolute.n.01 & 0.111111111111\\ \hline
materialism.n.02 & absolute.n.01 & 0.111111111111\\ \hline
strategy.n.02 & absolute.n.01 & 0.111111111111\\ \hline
idealism.n.01 & absolute.n.01 & 0.111111111111\\ \hline
imperialism.n.02 & people.n.01 & 0.111111111111\\ \hline
television receiver.n.01 & land.n.04 & 0.1\\ \hline
car.n.01 & land.n.04 & 0.1\\ \hline
theory.n.01 & people.n.01 & 0.1\\ \hline
sugar.n.01 & land.n.04 & 0.0909090909091\\ \hline
lion.n.01 & land.n.04 & 0.0714285714286\\ \hline
\multicolumn{3}{|c|}{Lenin: The Agrarian Programme  - lch similarity} \\ \hline
revolution.n.01 & revolution.n.01 & 3.63758615973\\ \hline
proletarian.n.01 & peasant.n.01 & 2.25129179861\\ \hline
money.n.01 & bill.n.03 & 2.02814824729\\ \hline
war.n.01 & nationalization.n.03 & 2.02814824729\\ \hline
practice.n.03 & nationalization.n.03 & 1.8458266905\\ \hline
capitalism.n.01 & people.n.01 & 1.8458266905\\ \hline
hair.n.01 & land.n.04 & 1.69167601067\\ \hline
toilet.n.04 & nationalization.n.03 & 1.44036158239\\ \hline
philosophy.n.02 & absolute.n.01 & 1.44036158239\\ \hline
materialism.n.02 & absolute.n.01 & 1.44036158239\\ \hline
strategy.n.02 & absolute.n.01 & 1.44036158239\\ \hline
idealism.n.01 & absolute.n.01 & 1.44036158239\\ \hline
imperialism.n.02 & people.n.01 & 1.44036158239\\ \hline
television receiver.n.01 & land.n.04 & 1.33500106673\\ \hline
car.n.01 & land.n.04 & 1.33500106673\\ \hline
theory.n.01 & people.n.01 & 1.33500106673\\ \hline
sugar.n.01 & land.n.04 & 1.23969088693\\ \hline
lion.n.01 & land.n.04 & 0.998528830111\\ \hline
\multicolumn{3}{|c|}{Lenin: The Agrarian Programme  - res similarity} \\ \hline
revolution.n.01 & revolution.n.01 & 10.3880624692\\ \hline
proletarian.n.01 & peasant.n.01 & 9.00601439892\\ \hline
money.n.01 & bill.n.03 & 6.2549318819\\ \hline
war.n.01 & nationalization.n.03 & 4.52385364841\\ \hline
philosophy.n.02 & absolute.n.01 & 3.29228159341\\ \hline
materialism.n.02 & absolute.n.01 & 3.29228159341\\ \hline
strategy.n.02 & absolute.n.01 & 3.29228159341\\ \hline
idealism.n.01 & absolute.n.01 & 3.29228159341\\ \hline
capitalism.n.01 & people.n.01 & 2.85529368615\\ \hline
imperialism.n.02 & absolute.n.01 & 2.64452085446\\ \hline
theory.n.01 & absolute.n.01 & 2.64452085446\\ \hline
toilet.n.04 & nationalization.n.03 & 2.33223775508\\ \hline
practice.n.03 & nationalization.n.03 & 2.33223775508\\ \hline
lion.n.01 & peasant.n.01 & 2.22415047123\\ \hline
television receiver.n.01 & peasant.n.01 & 1.53183374322\\ \hline
car.n.01 & peasant.n.01 & 1.53183374322\\ \hline
hair.n.01 & peasant.n.01 & 1.53183374322\\ \hline
sugar.n.01 & land.n.04 & 0.801759114954\\ \hline
\multicolumn{3}{|c|}{Lenin: The Agrarian Programme  - jcn similarity} \\ \hline
revolution.n.01 & revolution.n.01 & 1e+300\\ \hline
proletarian.n.01 & peasant.n.01 & 0.363995193783\\ \hline
money.n.01 & bill.n.03 & 0.30619633675\\ \hline
war.n.01 & farming.n.01 & 0.102813269009\\ \hline
car.n.01 & land.n.04 & 0.0903577501322\\ \hline
hair.n.01 & land.n.04 & 0.0884432572523\\ \hline
capitalism.n.01 & people.n.01 & 0.0880782305893\\ \hline
philosophy.n.02 & people.n.01 & 0.0801894085342\\ \hline
television receiver.n.01 & land.n.04 & 0.0742092466431\\ \hline
theory.n.01 & people.n.01 & 0.0738644334041\\ \hline
practice.n.03 & people.n.01 & 0.0728246934249\\ \hline
lion.n.01 & land.n.04 & 0.0689306688111\\ \hline
strategy.n.02 & people.n.01 & 0.0642953740963\\ \hline
sugar.n.01 & land.n.04 & 0.0640122327597\\ \hline
toilet.n.04 & people.n.01 & 0.0632469161145\\ \hline
materialism.n.02 & people.n.01 & 0.0598861536358\\ \hline
idealism.n.01 & people.n.01 & 0.056994188175\\ \hline
imperialism.n.02 & people.n.01 & 0.055706849963\\ \hline
\multicolumn{3}{|c|}{Lenin: The Agrarian Programme  - lin similarity} \\ \hline
revolution.n.01 & revolution.n.01 & 1.0\\ \hline
proletarian.n.01 & peasant.n.01 & 0.867659956556\\ \hline
money.n.01 & bill.n.03 & 0.792980998294\\ \hline
war.n.01 & farming.n.01 & 0.481925511595\\ \hline
philosophy.n.02 & absolute.n.01 & 0.341316252969\\ \hline
capitalism.n.01 & people.n.01 & 0.334654457361\\ \hline
strategy.n.02 & absolute.n.01 & 0.294289762784\\ \hline
materialism.n.02 & absolute.n.01 & 0.279961229282\\ \hline
idealism.n.01 & absolute.n.01 & 0.270226260066\\ \hline
theory.n.01 & absolute.n.01 & 0.259782259686\\ \hline
practice.n.03 & restoration.n.02 & 0.240021453557\\ \hline
toilet.n.04 & restoration.n.02 & 0.216821056563\\ \hline
imperialism.n.02 & absolute.n.01 & 0.213506139286\\ \hline
lion.n.01 & peasant.n.01 & 0.20131688425\\ \hline
car.n.01 & land.n.04 & 0.18905393848\\ \hline
hair.n.01 & land.n.04 & 0.185792500123\\ \hline
television receiver.n.01 & land.n.04 & 0.160696186575\\ \hline
sugar.n.01 & land.n.04 & 0.0930896185493\\ \hline
\end{tabular}
\end{center}