### 2014-03-02 paper/analysis_lenin_what_is_to_be_done?.tex

    \begin{center}
      \begin{tabular}{ | l | l | l |}
        \hline
        Choosen Sensens & Closest Top Sense & similarity  \\ \hline
\multicolumn{3}{|c|}{Lenin: What Is To Be Done? - path similarity} \\ \hline
imperialism.n.02 & political orientation.n.01 & 0.5\\ \hline
proletarian.n.01 & worker.n.01 & 0.25\\ \hline
war.n.01 & conflict.n.01 & 0.25\\ \hline
practice.n.03 & conflict.n.01 & 0.2\\ \hline
revolution.n.01 & conflict.n.01 & 0.166666666667\\ \hline
toilet.n.04 & trade.n.02 & 0.142857142857\\ \hline
money.n.01 & mass.n.01 & 0.142857142857\\ \hline
capitalism.n.01 & organization.n.01 & 0.142857142857\\ \hline
hair.n.01 & working.n.01 & 0.125\\ \hline
television receiver.n.01 & working.n.01 & 0.111111111111\\ \hline
car.n.01 & working.n.01 & 0.111111111111\\ \hline
philosophy.n.02 & political orientation.n.01 & 0.111111111111\\ \hline
materialism.n.02 & political orientation.n.01 & 0.111111111111\\ \hline
strategy.n.02 & political orientation.n.01 & 0.111111111111\\ \hline
idealism.n.01 & political orientation.n.01 & 0.111111111111\\ \hline
theory.n.01 & political orientation.n.01 & 0.111111111111\\ \hline
sugar.n.01 & local anesthetic.n.01 & 0.0909090909091\\ \hline
lion.n.01 & worker.n.01 & 0.0833333333333\\ \hline
\multicolumn{3}{|c|}{Lenin: What Is To Be Done? - lch similarity} \\ \hline
imperialism.n.02 & political orientation.n.01 & 2.94443897917\\ \hline
proletarian.n.01 & worker.n.01 & 2.25129179861\\ \hline
war.n.01 & conflict.n.01 & 2.25129179861\\ \hline
practice.n.03 & conflict.n.01 & 2.02814824729\\ \hline
revolution.n.01 & conflict.n.01 & 1.8458266905\\ \hline
toilet.n.04 & trade.n.02 & 1.69167601067\\ \hline
money.n.01 & mass.n.01 & 1.69167601067\\ \hline
capitalism.n.01 & organization.n.01 & 1.69167601067\\ \hline
hair.n.01 & working.n.01 & 1.55814461805\\ \hline
television receiver.n.01 & working.n.01 & 1.44036158239\\ \hline
car.n.01 & working.n.01 & 1.44036158239\\ \hline
philosophy.n.02 & political orientation.n.01 & 1.44036158239\\ \hline
materialism.n.02 & political orientation.n.01 & 1.44036158239\\ \hline
strategy.n.02 & political orientation.n.01 & 1.44036158239\\ \hline
idealism.n.01 & political orientation.n.01 & 1.44036158239\\ \hline
theory.n.01 & political orientation.n.01 & 1.44036158239\\ \hline
sugar.n.01 & local anesthetic.n.01 & 1.23969088693\\ \hline
lion.n.01 & worker.n.01 & 1.15267950994\\ \hline
\multicolumn{3}{|c|}{Lenin: What Is To Be Done? - res similarity} \\ \hline
imperialism.n.02 & political orientation.n.01 & 8.60196873746\\ \hline
sugar.n.01 & local anesthetic.n.01 & 4.61519673162\\ \hline
war.n.01 & conflict.n.01 & 4.52385364841\\ \hline
television receiver.n.01 & newspaper.n.01 & 3.25767902589\\ \hline
car.n.01 & newspaper.n.01 & 3.25767902589\\ \hline
toilet.n.04 & trade.n.02 & 3.14352023311\\ \hline
capitalism.n.01 & organization.n.01 & 2.85529368615\\ \hline
money.n.01 & mass.n.01 & 2.69110280536\\ \hline
philosophy.n.02 & political orientation.n.01 & 2.64452085446\\ \hline
materialism.n.02 & political orientation.n.01 & 2.64452085446\\ \hline
strategy.n.02 & political orientation.n.01 & 2.64452085446\\ \hline
idealism.n.01 & political orientation.n.01 & 2.64452085446\\ \hline
theory.n.01 & political orientation.n.01 & 2.64452085446\\ \hline
proletarian.n.01 & worker.n.01 & 2.33354524374\\ \hline
practice.n.03 & trade.n.02 & 2.33223775508\\ \hline
lion.n.01 & worker.n.01 & 2.22415047123\\ \hline
revolution.n.01 & trade.n.02 & 2.03632771779\\ \hline
hair.n.01 & worker.n.01 & 1.53183374322\\ \hline
\multicolumn{3}{|c|}{Lenin: What Is To Be Done? - jcn similarity} \\ \hline
imperialism.n.02 & political orientation.n.01 & 0.193405491818\\ \hline
war.n.01 & conflict.n.01 & 0.152829762599\\ \hline
car.n.01 & worker.n.01 & 0.110550521246\\ \hline
hair.n.01 & worker.n.01 & 0.107698235698\\ \hline
proletarian.n.01 & worker.n.01 & 0.104758893715\\ \hline
capitalism.n.01 & organization.n.01 & 0.101121795595\\ \hline
sugar.n.01 & local anesthetic.n.01 & 0.0929714058017\\ \hline
money.n.01 & organization.n.01 & 0.0929667353193\\ \hline
philosophy.n.02 & organization.n.01 & 0.0908595602215\\ \hline
lion.n.01 & worker.n.01 & 0.0900802228812\\ \hline
practice.n.03 & motion.n.03 & 0.0882827134582\\ \hline
television receiver.n.01 & worker.n.01 & 0.0873062903722\\ \hline
theory.n.01 & organization.n.01 & 0.0828237113739\\ \hline
revolution.n.01 & motion.n.03 & 0.0784056470694\\ \hline
toilet.n.04 & motion.n.03 & 0.0745896410117\\ \hline
strategy.n.02 & organization.n.01 & 0.0709786674118\\ \hline
materialism.n.02 & organization.n.01 & 0.0656431911358\\ \hline
idealism.n.01 & organization.n.01 & 0.0621845289629\\ \hline
\multicolumn{3}{|c|}{Lenin: What Is To Be Done? - lin similarity} \\ \hline
imperialism.n.02 & political orientation.n.01 & 0.768910941363\\ \hline
war.n.01 & conflict.n.01 & 0.580318438615\\ \hline
sugar.n.01 & local anesthetic.n.01 & 0.461833980698\\ \hline
car.n.01 & newspaper.n.01 & 0.411156952767\\ \hline
capitalism.n.01 & organization.n.01 & 0.366071452787\\ \hline
television receiver.n.01 & newspaper.n.01 & 0.356914166707\\ \hline
money.n.01 & mass.n.01 & 0.331362497131\\ \hline
proletarian.n.01 & worker.n.01 & 0.328371898491\\ \hline
philosophy.n.02 & political orientation.n.01 & 0.31307622756\\ \hline
toilet.n.04 & trade.n.02 & 0.312879932381\\ \hline
theory.n.01 & political orientation.n.01 & 0.294463460062\\ \hline
practice.n.03 & motion.n.03 & 0.291680639247\\ \hline
lion.n.01 & worker.n.01 & 0.286073258452\\ \hline
strategy.n.02 & political orientation.n.01 & 0.264762834975\\ \hline
materialism.n.02 & political orientation.n.01 & 0.25040844726\\ \hline
hair.n.01 & worker.n.01 & 0.248092928538\\ \hline
revolution.n.01 & motion.n.03 & 0.242033306584\\ \hline
idealism.n.01 & political orientation.n.01 & 0.240750703939\\ \hline
\end{tabular}
\end{center}