### 2014-03-02 paper/analysis_the_state_and_revolution.tex

    \begin{center}
      \begin{tabular}{ | l | l | l |}
        \hline
        Choosen Sensens & Closest Top Sense & similarity  \\ \hline
\multicolumn{3}{|c|}{The State and Revolution - path similarity} \\ \hline
revolution.n.01 & revolution.n.01 & 1.0\\ \hline
imperialism.n.02 & communism.n.02 & 0.333333333333\\ \hline
proletarian.n.01 & peer.n.01 & 0.25\\ \hline
practice.n.03 & work.n.01 & 0.2\\ \hline
toilet.n.04 & work.n.01 & 0.166666666667\\ \hline
war.n.01 & work.n.01 & 0.166666666667\\ \hline
money.n.01 & phase.n.01 & 0.125\\ \hline
philosophy.n.02 & law.n.04 & 0.125\\ \hline
materialism.n.02 & law.n.04 & 0.125\\ \hline
strategy.n.02 & law.n.04 & 0.125\\ \hline
idealism.n.01 & law.n.04 & 0.125\\ \hline
capitalism.n.01 & democracy.n.02 & 0.125\\ \hline
hair.n.01 & peer.n.01 & 0.111111111111\\ \hline
theory.n.01 & experience.n.03 & 0.1\\ \hline
television receiver.n.01 & peer.n.01 & 0.0833333333333\\ \hline
car.n.01 & peer.n.01 & 0.0833333333333\\ \hline
lion.n.01 & peer.n.01 & 0.0833333333333\\ \hline
sugar.n.01 & peer.n.01 & 0.0833333333333\\ \hline
\multicolumn{3}{|c|}{The State and Revolution - lch similarity} \\ \hline
revolution.n.01 & revolution.n.01 & 3.63758615973\\ \hline
imperialism.n.02 & communism.n.02 & 2.53897387106\\ \hline
proletarian.n.01 & peer.n.01 & 2.25129179861\\ \hline
practice.n.03 & work.n.01 & 2.02814824729\\ \hline
toilet.n.04 & work.n.01 & 1.8458266905\\ \hline
war.n.01 & work.n.01 & 1.8458266905\\ \hline
money.n.01 & phase.n.01 & 1.55814461805\\ \hline
philosophy.n.02 & law.n.04 & 1.55814461805\\ \hline
materialism.n.02 & law.n.04 & 1.55814461805\\ \hline
strategy.n.02 & law.n.04 & 1.55814461805\\ \hline
idealism.n.01 & law.n.04 & 1.55814461805\\ \hline
capitalism.n.01 & democracy.n.02 & 1.55814461805\\ \hline
hair.n.01 & peer.n.01 & 1.44036158239\\ \hline
theory.n.01 & experience.n.03 & 1.33500106673\\ \hline
television receiver.n.01 & peer.n.01 & 1.15267950994\\ \hline
car.n.01 & peer.n.01 & 1.15267950994\\ \hline
lion.n.01 & peer.n.01 & 1.15267950994\\ \hline
sugar.n.01 & peer.n.01 & 1.15267950994\\ \hline
\multicolumn{3}{|c|}{The State and Revolution - res similarity} \\ \hline
revolution.n.01 & revolution.n.01 & 10.3880624692\\ \hline
imperialism.n.02 & communism.n.02 & 8.60196873746\\ \hline
philosophy.n.02 & law.n.04 & 3.29228159341\\ \hline
materialism.n.02 & law.n.04 & 3.29228159341\\ \hline
strategy.n.02 & law.n.04 & 3.29228159341\\ \hline
idealism.n.01 & law.n.04 & 3.29228159341\\ \hline
toilet.n.04 & work.n.01 & 3.14352023311\\ \hline
capitalism.n.01 & club.n.02 & 2.85529368615\\ \hline
money.n.01 & phase.n.01 & 2.69110280536\\ \hline
theory.n.01 & law.n.04 & 2.64452085446\\ \hline
proletarian.n.01 & engels.n.01 & 2.33354524374\\ \hline
war.n.01 & work.n.01 & 2.33223775508\\ \hline
practice.n.03 & work.n.01 & 2.33223775508\\ \hline
lion.n.01 & engels.n.01 & 2.22415047123\\ \hline
television receiver.n.01 & engels.n.01 & 1.53183374322\\ \hline
car.n.01 & engels.n.01 & 1.53183374322\\ \hline
hair.n.01 & engels.n.01 & 1.53183374322\\ \hline
sugar.n.01 & engels.n.01 & 0.801759114954\\ \hline
\multicolumn{3}{|c|}{The State and Revolution - jcn similarity} \\ \hline
revolution.n.01 & revolution.n.01 & 1e+300\\ \hline
imperialism.n.02 & communism.n.02 & 0.14975874591\\ \hline
war.n.01 & work.n.01 & 0.119881984187\\ \hline
philosophy.n.02 & law.n.04 & 0.103411969458\\ \hline
practice.n.03 & work.n.01 & 0.103154641506\\ \hline
toilet.n.04 & work.n.01 & 0.0985118809398\\ \hline
money.n.01 & phase.n.01 & 0.0916158054473\\ \hline
theory.n.01 & work.n.01 & 0.0910758552239\\ \hline
car.n.01 & peer.n.01 & 0.0887294617483\\ \hline
hair.n.01 & peer.n.01 & 0.0868826421077\\ \hline
proletarian.n.01 & peer.n.01 & 0.0849595683478\\ \hline
strategy.n.02 & law.n.04 & 0.07841412673\\ \hline
lion.n.01 & peer.n.01 & 0.0750424463326\\ \hline
television receiver.n.01 & peer.n.01 & 0.0731074093034\\ \hline
capitalism.n.01 & club.n.02 & 0.0729022081709\\ \hline
materialism.n.02 & law.n.04 & 0.0719531280061\\ \hline
idealism.n.01 & law.n.04 & 0.0678185221973\\ \hline
sugar.n.01 & work.n.01 & 0.0625959407953\\ \hline
\multicolumn{3}{|c|}{The State and Revolution - lin similarity} \\ \hline
revolution.n.01 & revolution.n.01 & 1.0\\ \hline
imperialism.n.02 & communism.n.02 & 0.720392353354\\ \hline
philosophy.n.02 & law.n.04 & 0.405088626962\\ \hline
toilet.n.04 & work.n.01 & 0.382467580972\\ \hline
war.n.01 & work.n.01 & 0.35863993877\\ \hline
strategy.n.02 & law.n.04 & 0.340509805473\\ \hline
money.n.01 & phase.n.01 & 0.330250297796\\ \hline
practice.n.03 & work.n.01 & 0.324854541172\\ \hline
materialism.n.02 & law.n.04 & 0.321472637881\\ \hline
idealism.n.01 & law.n.04 & 0.308702564543\\ \hline
theory.n.01 & law.n.04 & 0.305328480717\\ \hline
capitalism.n.01 & club.n.02 & 0.29394209418\\ \hline
proletarian.n.01 & peer.n.01 & 0.28393127113\\ \hline
lion.n.01 & peer.n.01 & 0.250268807542\\ \hline
car.n.01 & peer.n.01 & 0.213736072979\\ \hline
hair.n.01 & peer.n.01 & 0.210222579299\\ \hline
television receiver.n.01 & peer.n.01 & 0.182991045418\\ \hline
sugar.n.01 & peer.n.01 & 0.0895207954547\\ \hline
\end{tabular}
\end{center}