### 2014-03-15 paper/handpicked_table.tex
fix some spelling

\begin{tabular}{ | L{2cm} | L{5cm} |}
    \hline
    Synset & Gloss \\ \hline
    \multicolumn{2}{|c|}{Human handpicked concepts to compare against top senses} \\ \hline
            war & the waging of armed conflict against an enemy \\ \hline
            capitalism & an economic system based on private ownership of capital\\ \hline
            theory & a well-substantiated explanation of some aspect of the natural world \\ \hline
            toilet & the act of dressing and preparing yourself \\ \hline
            imperialism & a political orientation that advocates imperial interests \\ \hline
            proletarian & a member of the working class (not necessarily employed) \\ \hline
            practice & translating an idea into action \\ \hline
            revolution & a drastic and far-reaching change in ways of thinking and behaving \\ \hline
            money & the most common medium of exchange; functions as legal tender \\ \hline
            philosophy & the rational investigation of questions about existence and knowledge and ethics \\ \hline
            idealism & (philosophy) the philosophical theory that ideas are the only reality \\ \hline
            strategy & the branch of military science dealing with military command and the planning and conduct of a war \\ \hline
            materialism & (philosophy) the philosophical theory that matter is the only reality \\ \hline
            lion & large gregarious predatory feline of Africa and India having a tawny coat with a shaggy mane in the male \\ \hline
            television\_receiver. & an electronic device that receives television signals and displays them on a screen \\ \hline
            hair & a covering for the body (or parts of it) consisting of a dense growth of threadlike structures (as on the human head); helps to prevent heat loss \\ \hline
            sugar & a white crystalline carbohydrate used as a sweetener and preservative \\ \hline
            car & a motor vehicle with four wheels; usually propelled by an internal combustion engine \\ \hline
\end{tabular}