### 2014-03-15 paper/results.tex

\begin{tabular}{p{1.1cm} p{1.1cm} p{1.1cm} p{1.5cm} p{1.5cm}}
        \toprule
        Work 1 & Work 2 & Work 3 & Work 4 & Work 5 \\
%        \multicolumn{2}{c}{Work 1} & \multicolumn{2}{c}{Work 2} & \multicolumn{2}{c}{Work 3} & \multicolumn{2}{c}{Work 4} & \multicolumn{2}{c}{Work 5} \\
        \midrule
        marx & enterprise & physics & local\_anesthetic & rent \\
        revolution & depository\_financial\_institution & sociable & organization  & land \\
        plant & imperialism & space & marriage & peasant \\
        engels & universe & time  & criticism & capital \\
        club & million & berkeley & revolutionist & revolution \\
        power & colony & truth & freedom & marx \\
        party & colonial &  aim & newspaper & people \\
        sociable &  product & sensation & worker & nationalization \\
        experience & penny & experience & consciousness & restoration \\
        bernstein & company  & engels & newspaper & right \\
        question & capitalist & motion & businessperson & private  \\
        equality & monopoly & marx & mass & farming \\
        phase & area &  book & motion & free \\
        commune & policy &  plant & german  & central \\
        one &  power & general & goverment & land \\
        democracy & banks &  helmholtz & political\_orientation & capitalism \\
        class &  worker & neccesity & party & democracy \\
        law &  imperialist & law & secret & landlord \\
        communism & company & sensation & working & farming \\
        peer & boards & energy & class & agency \\
        \bottomrule
    \end{tabular}
\newline
%aWork1 : The State and Revolution
%work2: Lenin: Imperialism, the Highest Stage of Capitalism
%Work3: Lenin: MATERIALISM and EMPIRIO-CRITICISM
%Work4: Lenin: What Is To Be Done?
%wotk5: Lenin: The Agrarian Programme of Social-Democracy in the First Russian Revolution, 1905-1907

    \begin{tabular}{ p{2cm} p{2cm} p{3cm}}
    \toprule
    Text &  Top Sense & Sense Detected \\
    \midrule
The Agrarian Programme & marx & United States comedian; one of four brothers who made motion pictures together (1901-1979) \\
    \midrule
The State and Revolution & bernstein & United States conductor and composer (1918-1990) \\
    \bottomrule
\end{tabular}

\input{top_senses_definitions}