### 2014-03-19 paper/top_senses_definitions.tex
fix top 5 senses table!

\begin{table*}[ht]
\caption{Top 5 Senses with Closest Handpicked Synset}
\centering
\begin{tabular}{ p{2.5cm} p{9cm} p{2.5cm}}
    \toprule
    Synset & Gloss  & Handpicked \cr
    \hline
    \midrule  \multicolumn{3}{c}{The State and Revolution} \\  \midrule
    revolution.n.01 & a drastic and far-reaching change in ways of thinking and behaving & revolution.n.01\\
    imperialism.n.02 & a political orientation that advocates imperial interests & communism.n.02\\
    proletarian.n.01 & a member of the working class (not necessarily employed) & peer.n.01\\
    war.n.01 & the waging of armed conflict against an enemy & experience.n.03\\
    practice.n.03 & translating an idea into action & experience.n.03\\
    \midrule  \multicolumn{3}{c}{Lenin: Imperialism, the Highest Stage of Capitalism} \\  \midrule
    proletarian.n.01 & a member of the working class (not necessarily employed) & capitalist.n.02\\
    hair.n.01 & a covering for the body (or parts of it) consisting of a dense growth of threadlike structures (as on the human head); helps to prevent heat loss & universe.n.01\\
    money.n.01 & the most common medium of exchange; functions as legal tender & penny.n.02\\
    practice.n.03 & translating an idea into action & monopoly.n.01\\
    toilet.n.04 & the act of dressing and preparing yourself & monopoly.n.01\\
    \midrule  \multicolumn{3}{c}{Lenin: MATERIALISM and EMPIRIO-CRITICISM} \\  \midrule
    revolution.n.01 & a drastic and far-reaching change in ways of thinking and behaving & experience.n.03\\
    proletarian.n.01 & a member of the working class (not necessarily employed) & helmholtz.n.01\\
    philosophy.n.02 & the rational investigation of questions about existence and knowledge and ethics & physics.n.01\\
    war.n.01 & the waging of armed conflict against an enemy & experience.n.03\\
    strategy.n.02 & the branch of military science dealing with military command and the planning and conduct of a war & physics.n.01\\
    \midrule  \multicolumn{3}{c}{Lenin: What Is To Be Done?} \\  \midrule
    imperialism.n.02 & a political orientation that advocates imperial interests & political\_orientation.n.01\\
    proletarian.n.01 & a member of the working class (not necessarily employed) & worker.n.01\\
    practice.n.03 & translating an idea into action & motion.n.03\\
    money.n.01 & the most common medium of exchange; functions as legal tender & mass.n.01\\
    war.n.01 & the waging of armed conflict against an enemy & motion.n.03\\

    \midrule  \multicolumn{3}{c}{Lenin: The Agrarian Programme of Social-Democracy in the First Russian Revolution, 1905-1907} \\  \midrule
    revolution.n.01 & a drastic and far-reaching change in ways of thinking and behaving & revolution.n.01\\
    capitalism.n.01 & an economic system based on private ownership of capital & capitalism.n.01\\
    proletarian.n.01 & a member of the working class (not necessarily employed) & peasant.n.01\\
    war.n.01 & the waging of armed conflict against an enemy & nationalization.n.03\\
    practice.n.03 & translating an idea into action & nationalization.n.03\\

    \bottomrule
\end{tabular}
\end{table*}