### 2014-03-19 paper/handpicked_table.tex
reorder handpicked_table and correct some typos

\begin{table*}[ht]
\caption*{Handpicked concepts to compare against top senses}
\centering
%\begin{tabular}{p{0.25\linewidth}p{0.25\linewidth}p{0.25\linewidth}}
\begin{tabular}{ p{2.5cm} p{15cm} }
    \toprule
    Synset & Gloss \\
    \midrule
            war & the waging of armed conflict against an enemy \\
            strategy & the branch of military science dealing with military command and the planning and conduct of a war \\
            practice & translating an idea into action \\
            money & the most common medium of exchange; functions as legal tender \\
            imperialism & a political orientation that advocates imperial interests \\
            proletarian & a member of the working class (not necessarily employed) \\
            revolution & a drastic and far-reaching change in ways of thinking and behaving \\
            capitalism & an economic system based on private ownership of capital\\
            materialism & (philosophy) the philosophical theory that matter is the only reality \\
            theory & a well-substantiated explanation of some aspect of the natural world \\
            philosophy & the rational investigation of questions about existence and knowledge and ethics \\
            idealism & (philosophy) the philosophical theory that ideas are the only reality \\
            lion & large gregarious predatory f of Africa and India having a tawny coat with a shaggy mane in the male \\
            toilet & the act of dressing and preparing yourself \\
            television\_receiver & an electronic device that receives television signals and displays them on a screen \\
            hair & a covering for the body (or parts of it) consisting of a dense growth of threadlike structures (as on the human head); helps to prevent heat loss \\
            sugar & a white crystal carbohydrate used as a sweetener and preservative \\
            car & a motor vehicle with four wheels; usually propelled by an internal combustion engine \\

    \bottomrule
\end{tabular}

\end{table*}