### 2014-03-22 paper/results.tex
remove everything but page number from footer

\begin{table*}[ht]
\caption*{Top sense for all Lenin works}
\centering
\begin{tabular}{p{3cm} p{3cm} p{3cm} p{3cm} p{3cm}}
        \toprule
        The State and Revolution & Imperialism,the Highest .. & Materialism and ... & What Is To Be Done? &  The Agrarian Programme ... \\
%        \multicolumn{2}{c}{Work 1} & \multicolumn{2}{c}{Work 2} & \multicolumn{2}{c}{Work 3} & \multicolumn{2}{c}{Work 4} & \multicolumn{2}{c}{Work 5} \\
        \midrule
        marx & enterprise & physics & local\_anesthetic & rent \\
        revolution & depository\_financial\_institution & sociable & organization  & land \\
        plant & imperialism & space & marriage & peasant \\
        engels & universe & time  & criticism & capital \\
        club & million & berkeley & revolutionist & revolution \\
        power & colony & truth & freedom & marx \\
        party & colonial &  aim & newspaper & people \\
        sociable &  product & sensation & worker & nationalization \\
        experience & penny & experience & consciousness & restoration \\
        bernstein & company  & engels & newspaper & right \\
        question & capitalist & motion & businessperson & private  \\
        equality & monopoly & marx & mass & farming \\
        phase & area &  book & motion & free \\
        commune & policy &  plant & german  & central \\
        one &  power & general & goverment & land \\
        democracy & banks &  helmholtz & political\_orientation & capitalism \\
        class &  worker & neccesity & party & democracy \\
        law &  imperialist & law & secret & landlord \\
        communism & company & sensation & working & farming \\
        peer & boards & energy & class & agency \\
        \bottomrule
        \multicolumn{5}{p{\linewidth}}{The table shows $top-senses$ obtained from the choosen works. Each $top-sense$ is ordered according to the information value score that we obtained previous to apply the $WSD$ algorithm. Most of the senses found by our algorithm were the senses related to the works, except for some of them. On Whats is to be done? local\_anesthetic was matched as sense. On State and revolution Marx was matched with a United States comedian and the same result we got with Berstein that was matched with a composer} \\
    \end{tabular}
\end{table*}